---
title: "R Notebook"
output: html_notebook
---

#### 1.1 load data
```{r}
library(scATFR)
library(PIDC)
exp_data <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/EHT_gene_count.txt",sep="\t",header=T,row.names=1)
gene_infor <- read.table("D:/HSC_Autophagy/ref_genome/gencode.vM25.gene.infor.rm.version.tsv",sep="\t",header=T,row.names=1)
col_metadata <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/sample2stage.txt",header=T,sep="\t")
row.names(col_metadata) <- col_metadata$sample
exp_data_median <- apply(exp_data,2,function(x){tapply(x,as.character(gene_infor[names(x),"gene_name"]),sum)})
exp_data_median <- exp_data_median[apply(exp_data_median,1,function(x){length(x[x>0])>=10}),]
saveRDS(exp_data_median, "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/EHT_gene_count.rds")

atfr <- scATFR(exp_data = exp_data_median,col_data = col_metadata)
atfr <- scater::logNormCounts(atfr)
saveRDS(atfr,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/GRNs/atfr_init.rds")
```

#### 1.2 perform GRN infering

```{r}
atfr <- inferGRNs(x = atfr,method="pidc",use_assay="logcounts", ncores=6)
saveRDS(GRN(atfr,"pidc"),"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/GRNs/pidc_GRN.rds")
pidc_nets <- matToNet(weightMat = GRN(atfr,"pidc"),cutoff = 0)
saveRDS(pidc_nets,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/GRNs/pidc_regulons.rds")

```

#### 1.3 integrate GRNs 

```{r}
mouse_tfs <- intersect(TF_Symbols$mouse,row.names(atfr))
blood_nets <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/final_network/mouse/blood_merged_filter_strict_network.txt.gz",sep="\t",header=T) %>% filter(tf %in% mouse_tfs)
pidc_nets <- pidc_nets %>% filter(regulator %in% mouse_tfs)
regulon_nets <- integraNets(grnNets = list(pidc_nets, blood_nets), norm_weight = T,
                            max_target_num=2000, min_target_num = 100, 
                            join_type = "full", return_sig = T, maxInter = 5, 
                            ncores = 6)
                        
interged_regulons <- dfToList(regulon_nets)
length(interged_regulons)
```

#### 1.3 filtering and calculating active GRNs 

```{r}
atfr <- filterRegulons(x = atfr, gene_list = interged_regulons, maxSize=500, minSize=5)
atfr <- regulonActivity(x = atfr, method="viper", use_assay="logcounts")
active_mat <- assay(altExp(atfr,"viper"))
write.table(active_mat,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/tf_regulon_active_mat.txt",sep="\t",quote=F)

atfr <- regulonActivity(x = atfr, method="aucell", use_assay="logcounts")
active_mat2 <- assay(altExp(atfr,"aucell"))
write.table(active_mat2,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/tf_regulon_active_mat_auell.txt",sep="\t",quote=F)

atfr <- regulonActivity(x = atfr, method="ssgsea", use_assay="logcounts")
active_mat3 <- assay(altExp(atfr,"ssgsea"))
write.table(active_mat3,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/tf_regulon_active_mat_ssgsea.txt",sep="\t",quote=F)
saveRDS(atfr,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")

```

#### 1.4 active mat generate from dorothea

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
tf_gene <- intersect(TF_Symbols$mouse, row.names(atfr))
library(dorothea)
regulons <- dorothea_mm %>% filter(tf %in% row.names(atfr))

exprMatr <- as.matrix(assay(atfr,"logcounts"))
tf_activities <- run_viper(exprMatr, regulons, options =  list(method = "scale", minsize = 4, 
                                                  eset.filter = FALSE, cores = 1, verbose = TRUE))
act_sce <- SingleCellExperiment::SingleCellExperiment(list(active_score=tf_activities))
colData(act_sce) <- colData(atfr)
colData(act_sce)$stage <- reLevel(vect = colData(act_sce)$stage)
act_sce <- scater::runPCA(x=act_sce, exprs_values="active_score") 
plotReducedDim(act_sce, dimred="PCA",colour_by = "stage")
act_sce <- scater::runTSNE(x = act_sce,exprs_values="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "stage")
#act_sce <- scater::runUMAP(x = act_sce, subset_row=tf_gene,exprs_values="active_score")
#plotReducedDim(act_sce, dimred="UMAP",colour_by = "stage")


```




### 2 expression level analysis

#### 2.1 expression level quality control

```{r}
exp_data_median <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/EHT_gene_count.rds")
col_metadata <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/sample2stage.txt",header=T,sep="\t")
row.names(col_metadata) <- col_metadata$sample
gene_number <- apply(exp_data_median,2,function(x){length(x[x>=1])})
col_metadata$gene_number <- gene_number[row.names(col_metadata)]
col_metadata$stage <- factor(col_metadata$stage,levels=unique(col_metadata$stage))
stage2color <- read.table("D:/HSC_Autophagy/2.HSC_formation/sample_infor/stage2color.txt",sep="\t",header=T,comment.char = "")
row.names(stage2color) <- stage2color$stage

text_size = 8
line_size = 0.5/1.07
library(ggplot2)
#------set plot theme
mytheme = theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA,color=NA),
strip.background = element_rect(fill=NA,color=NA),
strip.text = element_text(size=text_size,color="black"),
panel.grid.major = element_line(size=line_size/2, linetype = 8, color = "lightgray"),
panel.grid.minor = element_line(size=NA, color = NA),
axis.ticks = element_line(size=line_size,color="black"),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45, hjust = 1, vjust =1,color="black"), 
axis.text.y= element_text(size=text_size,color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

plot_frm <- col_metadata

p <- ggplot(plot_frm,mapping = aes(x=stage,y=gene_number, color=stage))+
  geom_violin(position = "dodge", na.rm = TRUE, scale = "width",width = 0.9)+
  geom_boxplot(width=0.25,fill=NA, size=0.2,outlier.colour=NA, alpha = 0.8)+
  geom_jitter(width=0.2,size=0.2,shape=16,color="black", alpha = 1)+
  scale_color_manual(values = stage2color$figurecolor)+
  labs(x="",y='number of expressed genes in each single cells')+
  mytheme

p

ggsave(filename ="EHT_gene_num_violin.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/quality_control/figure",
       width = 80, height =50, dpi = 600, limitsize = FALSE, units ="mm")
```

#### 2.2 dimensionality reduction using expression data and activate score

```{r}
library(scater)
library(scran)
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
tf_gene <- intersect(TF_Symbols$mouse,row.names(atfr))
exp_sce <- atfr[tf_gene,]
exp_sce <- scater::logNormCounts(exp_sce)
colData(exp_sce) <- colData(atfr)
colData(exp_sce)$stage <- reLevel(vect = colData(exp_sce)$stage)

log_exp <- as.matrix(logcounts(exp_sce))
log_exp_sd <- apply(log_exp,1,sd)
log_exp_sd_sort <- sort(log_exp_sd,decreasing = T)
log_exp_sd_top <- names(head(log_exp_sd_sort,500))

exp_sce <- scater::runPCA(x=exp_sce,subset_row=log_exp_sd_top,exprs_values = "logcounts") 
plotReducedDim(exp_sce, dimred="PCA",colour_by = "stage")
exp_sce <- scater::runTSNE(x = exp_sce,subset_row=log_exp_sd_top,exprs_values = "logcounts")
p <- plotReducedDim(exp_sce, dimred="TSNE",colour_by = "stage",point_size=0.5,point_alpha=1)
#exp_sce  <- scater::runUMAP(x = exp_sce, subset_row=tf_gene)
#plotReducedDim(exp_sce , dimred="UMAP",colour_by = "stage")
saveRDS(exp_sce,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.1_TF_exp_sce.rds")

#--- active score
act_sce <- SingleCellExperiment::SingleCellExperiment(list(active_score=assay(altExp(atfr,"viper"))))
colData(act_sce) <- colData(atfr)
colData(act_sce)$stage <- reLevel(vect = colData(act_sce)$stage)
act_mat <- as.matrix(assay(act_sce,"active_score"))
act_mat_sd <- apply(act_mat,1,sd)
act_mat_sd_sort <- sort(act_mat_sd,decreasing = T)
act_mat_cv_top <- names(head(act_mat_cv_sort,500))

act_sce <- scater::runPCA(x=act_sce,subset_row=act_mat_cv_top, exprs_values="active_score") 
plotReducedDim(act_sce, dimred="PCA",colour_by = "stage")
act_sce <- scater::runTSNE(x = act_sce,subset_row=act_mat_cv_top, exprs_values="active_score")
q <- plotReducedDim(act_sce, dimred="TSNE",colour_by = "stage",point_size=0.5,point_alpha=1)
#act_sce <- scater::runUMAP(x = act_sce, subset_row=tf_gene,exprs_values="active_score")
#plotReducedDim(act_sce, dimred="UMAP",colour_by = "stage")
saveRDS(act_sce,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")

p1 <- plotReducedDim(exp_sce, dimred="TSNE",colour_by = "stage",point_size=0.5,point_alpha=1)
p2 <- plotReducedDim(act_sce, dimred="TSNE",colour_by = "stage",point_size=0.5,point_alpha=1)

p <- p1+p2

ggsave(filename ="tf_and_tf_regulons_tSNE_plot.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/figure",
       width = 160, height =50, dpi = 600, limitsize = FALSE, units ="mm")


#---all gene
colData(atfr)$stage <- reLevel(vect = colData(atfr)$stage)
log_exp_all <- as.matrix(logcounts(atfr))
log_exp_all_cv <- apply(log_exp_all,1,sd)
log_exp_cv_all_sort <- sort(log_exp_all_cv,decreasing = T)
log_exp_cv_all_top <- names(head(log_exp_cv_all_sort,500))
atfr_sce <- scater::runPCA(x=atfr, subset_row=log_exp_cv_all_top, exprs_values="logcounts") 
plotReducedDim(atfr_sce, dimred="PCA",colour_by = "stage")
atfr_sce <- scater::runTSNE(x = atfr_sce,subset_row=log_exp_cv_all_top, exprs_values="logcounts")
r <- plotReducedDim(atfr_sce, dimred="TSNE",colour_by = "stage",point_size=0.5,point_alpha=1)



ggsave(filename ="tSNE_tf_all_regulon_level.pdf" ,plot = p+r+q, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/figure/",
       width = 230, height =50, dpi = 600, limitsize = FALSE, units ="mm")

```



#### 2.3 TF expression vs regulon active matrix correlation

```{r}
exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")
assay(act_sce,"logcounts") <- assay(exp_sce,"logcounts")[row.names(assay(act_sce,"active_score")),]
acticate_mat <- assay(act_sce,"active_score")
exp_mat <- assay(exp_sce,"logcounts")[row.names(acticate_mat),]
cor_p <- sapply(row.names(acticate_mat),function(x,acticate_mat, exp_mat){
  cor(acticate_mat[x,],exp_mat[x,])
},acticate_mat=acticate_mat, exp_mat=exp_mat)
names(cor_p) <- row.names(acticate_mat)

others_cor <- sapply(row.names(acticate_mat),function(x,acticate_mat, exp_mat){
  exp_mat_filter <- exp_mat[setdiff(row.names(exp_mat),x),]
  apply(exp_mat_filter,1,function(y){cor(acticate_mat[x,],y)})
  },acticate_mat=acticate_mat, exp_mat=exp_mat)
others_cor_frm <- data.frame(cor_p=as.vector(others_cor),type=rep("control",length(cor_p)))
plot_frm <- data.frame(cor_p,type=rep("case",length(cor_p)))
plot_frm_all <- rbind(others_cor_frm, plot_frm)

library(ggplot2)
p <- ggplot(plot_frm_all,aes(x=cor_p,fill=type,group=type))+
  geom_density()+
  scale_x_continuous(limits = c(-1,1))+
   mytheme
p
ggsave(filename ="tf_and_tf_regulons_active_matrix_similarity.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/figure",
       width = 75, height =50, dpi = 600, limitsize = FALSE, units ="mm")

```
#### 2.4 differential test

##### 2.4.1 cell type specific TF gene and TF regulons at expression and activate level

```{r}
#---expression level
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
exp_data <- logcounts(atfr)
TF_exp_mat <- as.matrix(exp_data[intersect(TF_Symbols$mouse,row.names(exp_data)),])
col_metadata <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/sample2stage.txt",header=T,sep="\t")
row.names(col_metadata) <- col_metadata$sample

cts_tfs <- findAllCTSRegulons(x = TF_exp_mat, col_data= col_metadata, clust_name="stage", fdr = 1, css=0, nperm = 10000, ncores=6)
saveRDS(cts_tfs,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
cts_tfs_diff <- cts_tfs %>% filter(padj<0.01)
saveRDS(cts_tfs_diff,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_diff_res.rds")

#---regulons level

cts_tf_regulons <- findAllCTSRegulons(x = atfr, clust_name="stage", use_assay="aucell", fdr=1L, css=0,  expression_per=0,  nperm=10000, ncores=6)
saveRDS(cts_tf_regulons,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01)
saveRDS(cts_tf_regulons_diff,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_diff_res_aucell.rds")

```

##### 2.4.2 AEC vs HEC TF gene and TF regulons at expression and activate level

```{r}
exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")
#assay(act_sce,"logcounts") <- assay(exp_sce,"logcounts")[row.names(assay(act_sce,"active_score")),]
col_data <- colData(exp_sce)
library(scran)
exp_sce_AEC_HEC <- findMarkers(x = exp_sce,groups=col_data$stage, restrict=c("AEC","HEC"))$HEC
exp_sce_AEC_HEC <- as.data.frame(exp_sce_AEC_HEC)
exp_sce_AEC_HEC_diff <- exp_sce_AEC_HEC %>% dplyr::filter(FDR<0.05)
exp_sce_AEC_HEC_not_diff <- exp_sce_AEC_HEC %>% dplyr::filter(FDR>0.2)
exp_sce_AEC_HEC_up <- exp_sce_AEC_HEC_diff %>% filter(logFC.AEC>=1)
exp_sce_AEC_HEC_down <- exp_sce_AEC_HEC_diff %>% filter(logFC.AEC<= -1)

assay(act_sce) <- t(apply(assay(act_sce),1,function(x){10*(x-min(x))/(max(x)-min(x))}))

act_sce_AEC_HEC <- findMarkers(x = act_sce,groups=col_data$stage, restrict=c("AEC","HEC"), assay.type = "active_score")$HEC
act_sce_AEC_HEC <- as.data.frame(act_sce_AEC_HEC)
act_sce_AEC_HEC_not_diff <- act_sce_AEC_HEC %>% dplyr::filter(FDR>0.2)
act_sce_AEC_HEC_diff <- act_sce_AEC_HEC %>% dplyr::filter(FDR<0.05)
act_sce_AEC_HEC_up <- act_sce_AEC_HEC_diff %>% filter(logFC.AEC>=1)
act_sce_AEC_HEC_down <- act_sce_AEC_HEC_diff %>% filter(logFC.AEC<= -1)

AEC_HEC_act_up_only <- intersect(row.names(act_sce_AEC_HEC_up), union(row.names(exp_sce_AEC_HEC_not_diff),row.names(exp_sce_AEC_HEC_down)))

AEC_HEC_frm <- data.frame(act_sce_AEC_HEC[AEC_HEC_act_up_only,],
                          exp_sce_AEC_HEC[AEC_HEC_act_up_only,])
AEC_HEC_frm_order <- AEC_HEC_frm[order(AEC_HEC_frm$FDR,AEC_HEC_frm$FDR,decreasing = c(F,T),method = "radix"),]
write.table(AEC_HEC_frm_order,"")
early_o <- intersect(earlyAEC_HEC_act_up_only,i)
late_o <- intersect(lateAEC_HEC_act_up_only,i)


i <- AEC_HEC_act_up_only
i <- c("Nfyc","Safb2","Fli1","Ash1l","Zkscan3","Kdm5b")
ncols <- round(sqrt(length(i)))+1
p1 <- plotExpression(object = exp_sce,features = i,x = "stage", colour_by = "stage", exprs_values = "logcounts", ncol = ncols, point_size=0.1)
p2 <- plotExpression(object = act_sce,features = i,x = "stage", colour_by = "stage",exprs_values = "active_score", ncol = ncols, point_size=0.1)
p1+p2

ggsave(filename ="tf_and_tf_regulons_active_matrix_similarity.pdf" ,plot = p1+p2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/figure",
       width = 1000, height =500, dpi = 600, limitsize = FALSE, units ="mm")
saveRDS(list(exp_sce_AEC_HEC, act_sce_AEC_HEC),"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/AEC_HEC/exp_act_diff_test.rds")

EHT_diff_res <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/AEC_HEC/exp_act_diff_test.rds")


```

#### 2.4.3 cts tf vs regulon distribution

```{r}
exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")
col_data <- as.data.frame(colData(exp_sce))
cts_tfs <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
cts_tfs_diff <- cts_tfs %>% filter(padj<0.01 & CSS >0.3)
cts_tfs_diff_list <- split(cts_tfs_diff,cts_tfs_diff$stage)

cts_tf_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01 & CSS >0.3 & exp_percent>0.3)
cts_tf_regulons_diff_list <- split(cts_tf_regulons_diff,cts_tf_regulons_diff$stage)

cts_tfs_regulon_comparison <- purrr::map2(.x = cts_tfs_diff_list, .y = cts_tf_regulons_diff_list, .f = function(x,y){
  overlap_gene <- intersect(x$regulon,y$regulon)
  tf_only <- setdiff(x$regulon,y$regulon)
  regulon_only <- setdiff(y$regulon,x$regulon)
  cts_gene_type <- c(rep("both",length(overlap_gene)), 
                   rep("tf_only",length(tf_only)),
                   rep("regulon_only",length(regulon_only)))
  names(cts_gene_type) <-c(overlap_gene, tf_only, regulon_only)
  x$type <- cts_gene_type[x$regulon]
  y$type <- cts_gene_type[y$regulon]
  list(TF=x,regulon=y)
})

cts_tfs_list <- lapply(cts_tfs_regulon_comparison,function(x){x[['TF']]})
cts_tfs_df <- do.call(rbind,cts_tfs_list)
write.table(cts_tfs_df,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/cts_tfs_level.tsv",sep="\t",quote=F,row.names=F)


cts_tf_regulon_list <- lapply(cts_tfs_regulon_comparison,function(x){x[['regulon']]})
cts_tf_regulon_df <- do.call(rbind, cts_tf_regulon_list)
fwrite(cts_tf_regulon_df,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/cts_tf_regulon_level.tsv",sep="\t",quote=F,row.names=F)
#---exp level
exp_mat <- as.matrix(logcounts(exp_sce))
cts_tfs_exp <- exp_mat[unique(cts_tfs_diff$regulon),]
cts_tfs_exp_mean <- t(apply(cts_tfs_exp,1,function(x){tapply(x,col_data[names(x),]$stage, mean)})) %>% as.data.frame()
cts_tfs_exp_mean$gene_name <- row.names(cts_tfs_exp_mean)

cts_tfs_df_plot <- cts_tfs_df[,c("stage","regulon","type")]
plot_tfs <- left_join(cts_tfs_df_plot, cts_tfs_exp_mean, by=c("regulon"="gene_name"))

stages <- c("AEC","HEC", "T1_pre_HSC", "T2_pre_HSC", "FL_HSC", "Adult_HSC")
plot_tfs$stage <- factor(plot_tfs$stage,levels=stages)
rv1 <- RadvizEnrichPlot(plot_tfs, anchors=stages, color_by="stage",
                       optim_anchor=F,return_rv=F,text_size = 5, point_size=1,alpha=0.6)+
  scale_color_manual(values = c('#729ece','#ff9e4a','#67bf5c','#ed665d','#ad8bc9','#a8786e'))
rv1

#---act level
act_mat <- as.matrix(assay(act_sce,"active_score"))
cts_act_mat <- act_mat[unique(cts_tf_regulons_diff$regulon),]
cts_act_mat <- t(apply(cts_act_mat,1,function(x){(x-min(x))/(max(x)-min(x))}))
cts_act_median <- t(apply(cts_act_mat,1,function(x){tapply(x,col_data[names(x),]$stage,mean)})) %>% as.data.frame()
cts_act_median$gene_name <- row.names(cts_act_median)

cts_tf_regulon_df_plot <- cts_tf_regulon_df[,c("stage","regulon","type")]
plot_act <- left_join(cts_tf_regulon_df_plot, cts_act_median, by=c("regulon"="gene_name"))

stages <- c("AEC","HEC", "T1_pre_HSC", "T2_pre_HSC", "FL_HSC", "Adult_HSC")
plot_act$stage <- factor(plot_act$stage,levels=stages)
rv2 <- RadvizEnrichPlot(plot_act, anchors=stages, color_by="stage",
                       optim_anchor=F, return_rv=F, text_size = 5, point_size=1,alpha=0.6)+
  scale_color_manual(values = c('#729ece','#ff9e4a','#67bf5c','#ed665d','#ad8bc9','#a8786e'))
rv2
ggsave(filename ="CTS_tf_and_regulon_Radviz.pdf" ,plot = rv1+rv2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/figure/",
       width = 200, height =100, dpi = 600, limitsize = FALSE, units ="mm")

rv1+rv2

tfs_regulon_overlap_gene <- cts_tfs_df %>% filter(type %in% "both")



```

#### 2.5 test using KO data

```{r}
Prdm16_ko <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/Prdm16.txt",sep="\t",header=T)
library(tidyr)
Prdm16_ko <- Prdm16_ko %>% separate_rows(Gene.symbol,sep = '///')
Prdm16_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Prdm16")

overlap_genes <- intersect(unlist(Prdm16_regulon_df$targets)[1:50],Prdm16_ko$Gene.symbol)

diff <- Prdm16_ko %>% filter(SYMBOL %in% overlap_genes)

Prdm16_ko_logFC <- tapply(Prdm16_ko$log2FC,Prdm16_ko$SYMBOL,mean)
Prdm16_ko_logFC <- Prdm16_ko_logFC[-1]

library(fgsea)
enrich <- fgsea(Prdm16_regulon_df$targets, Prdm16_ko_logFC)
p <- plotEnrichment(Prdm16_regulon_df$targets[["Prdm16"]],Prdm16_ko_logFC)

ggsave(filename ="Rela_KO_diff.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/figure",
       width = 150, height =100, dpi = 600, limitsize = FALSE, units ="mm")

```

#### 2.5.2 Atf4

```{r}
Atf4_ko <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE66380_RNAseq_gene_reads_count.txt.gz",sep="\t",header=T)
Atf4_ko$logFC <- log2((Atf4_ko$KO+1)/(Atf4_ko$WT+1))

Atf4_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Atf4")

overlap_genes <- intersect(unlist(Atf4_regulon_df$targets$Atf4),Atf4_ko$Gene)

Atf4_ko_logFC <- tapply(Atf4_ko$logFC,Atf4_ko$Gene,mean)

Atf4_enrich <- fgsea(Atf4_regulon_df$targets[1], Atf4_ko_logFC)


p <- plotEnrichment(Atf4_regulon_df$targets[[1]],Atf4_ko_logFC)


```

### 2.5.3 Sp1

```{r}
Sp1_ko <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE52497.Sp1.E14_pro.table.tsv",sep="\t",header=T)
Sp1_ko <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE52497.Sp1.HE1.table.tsv",sep="\t",header=T)

gene_infor <- fread("D:/HSC_Autophagy/important_genes/orthlogo_gene/NCBI/gene_info/gene_info",sep="\t",header=T)
gene_2_ensembl <- fread("D:/HSC_Autophagy/important_genes/orthlogo_gene/NCBI/gene2ensembl/gene2ensembl",sep="\t",header=T)
gene_2_ensembl <- gene_2_ensembl %>% filter(`#tax_id` == 10090)
gene_2_ensembl$RNA_nucleotide_accession.version <- sapply(str_split(gene_2_ensembl$RNA_nucleotide_accession.version,pattern = "\\."),function(x){x[1]})

mmu_gene_infor <- gene_infor %>% filter(`#tax_id` == 10090)
mmu_gene_infor <- full_join(mmu_gene_infor,gene_2_ensembl[,c("GeneID","RNA_nucleotide_accession.version")], by=c("GeneID"="GeneID"))
mmu_gene_infor<- mmu_gene_infor %>% select(Symbol,RNA_nucleotide_accession.version) %>% unique()

Sp1_ko <- left_join(x = Sp1_ko,y = mmu_gene_infor, by=c("GB_ACC"="RNA_nucleotide_accession.version"))

Sp1_ko_logFC <- tapply(Sp1_ko$logFC,Sp1_ko$Symbol,mean)

Sp1_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Sp1")

Sp1_enrich <- fgsea(Sp1_regulon_df$targets[1], -Sp1_ko_logFC)

p <- plotEnrichment(Sp1_regulon_df$targets[[1]], -Sp1_ko_logFC)

ggsave(filename ="Sp1_KO_diff.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/figure",
       width = 150, height =100, dpi = 600, limitsize = FALSE, units ="mm")
Sp1_ko %>% filter(Symbol=="Sp1")

```

#### 2.5.4 Foxm1

```{r}
Foxm1_ko <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE62360.Foxm1.table.tsv",sep="\t",header=T)
cts_tf_regulon_df <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
cts_tf_regulon_df <- cts_tf_regulon_df %>% filter(padj<0.01 & CSS>0.3 & exp_percent>0.3)

library(tidyr)
Foxm1_ko <- Foxm1_ko %>% separate_rows(Gene.symbol,sep = '///')
Foxm1_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Foxm1")

overlap_genes <- intersect(unlist(Foxm1_regulon_df$targets),Foxm1_ko$Gene.symbol)

diff <- Foxm1_ko %>% filter(Gene.symbol %in% overlap_genes)

Foxm1_ko_logFC <- tapply(Foxm1_ko$logFC,Foxm1_ko$Gene.symbol,mean)
Foxm1_ko_logFC <- Foxm1_ko_logFC[-1]

Foxm1_enrich <- fgsea(Foxm1_regulon_df$targets[1], Foxm1_ko_logFC)

p <- plotEnrichment(Foxm1_regulon_df$targets[[1]], Foxm1_ko_logFC)

```
#### 2.5.5 Spi1

```{r}
dir <- "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE65344_Spi1_RAW"
files <- list.files(dir)
fpkm_all <- lapply(files,function(x,dir=dir){
  in_FPKM <- read.table(file.path(dir,x),sep="\t",header=T)
  in_FPKM[,c('gene_short_name','q2_FPKM')]
},dir=dir)
names(fpkm_all) <- files
fpkm_all_tab1 <- fpkm_all[[1]]
fpkm_all_tab2 <- fpkm_all[[2]]
fpkm_all_tab3 <- fpkm_all[[3]]
fpkm_all_tab4 <- fpkm_all[[4]]
fpkm_all_tab5 <- fpkm_all[[5]]
fpkm_all_tab6 <- fpkm_all[[6]]

fpkm_all_tab1_2 <- left_join(fpkm_all_tab4,fpkm_all_tab6,by=c("gene_short_name"="gene_short_name"))
fpkm_all_tab1_2$logFC <- (fpkm_all_tab1_2$q2_FPKM.y)/(fpkm_all_tab1_2$q2_FPKM.x+1)
#fpkm_all_tab1_2$logFC <- log2(fpkm_all_tab1_2$logFC)
fpkm_all_tab1_logFC <- tapply(fpkm_all_tab1_2$logFC,fpkm_all_tab1_2$gene_short_name,mean)


Spi1_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Spi1")

Spi1_enrich <- fgsea(Spi1_regulon_df$targets[1], fpkm_all_tab1_logFC)

library(dplyr)
fpkm_all_tab <- purrr::reduce(.x = fpkm_all, .f = dplyr::left_join, by=c("gene_short_name"="gene_short_name"))


```




#### 2.5 cell type specific TF vs regulons CSS comparison using histogram

```{r}
cts_tfs <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
cts_tf_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")

text_size = 8
line_size = 0.5/1.07
library(ggplot2)
#------set plot theme
mytheme = theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA,color=NA),
strip.background = element_rect(fill=NA,color=NA),
strip.text = element_text(size=text_size,color="black"),
panel.grid.major = element_line(size=line_size/2, linetype = 8, color = "lightgray"),
panel.grid.minor = element_line(size=NA, color = NA),
axis.ticks = element_line(size=line_size,color="black"),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, hjust=0.5,vjust=0.5,color="black"), 
axis.text.y= element_text(size=text_size,color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

p <- ggplot()+
  geom_histogram(data = cts_tfs,mapping = aes(x=CSS,y=..count..),bins = 50,fill="#729ece",color=NA,alpha=0.7)+
  geom_histogram(data = cts_tf_regulons,mapping = aes(x=CSS,y=..count..),bins = 50,fill="#ed665d",color=NA,alpha=0.7)+
  scale_x_continuous(breaks=seq(0.0,0.6,0.2))+
  mytheme
p

ggsave(filename ="tf_and_tf_regulons_CSS_histgogram.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/figure",
       width = 60, height =50, dpi = 600, limitsize = FALSE, units ="mm")

```


### 3 cell type specific TF vs regulons comparison

#### 3.1 cell type specific TF vs regulons overlaps comparison

```{r}
library(data.table)
cts_tfs <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
fwrite(cts_tfs,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.tsv",sep="\t")
cts_tfs_diff <- cts_tfs %>% filter(padj<0.01 & CSS >0.3)
fwrite(cts_tfs_diff,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_diff_res.tsv",sep="\t")
cts_tfs_diff_stat <- table(cts_tfs_diff$regulon,cts_tfs_diff$stage)
attributes(cts_tfs_diff_stat)$class <- "matrix"
data_set <- as.data.frame(cts_tfs_diff_stat)
data_set <- data_set[,unique(colData(atfr)$stage)]

library(UpSetR)
p1 <- upset(data_set, nsets = 6, nintersects = 40, mb.ratio = c(0.5, 0.5),
      order.by = c("degree","freq"), decreasing = c(F,TRUE),point.size=5,
      line.size=2.5)
cts_tfs_diff_type <- tapply(cts_tfs_diff$stage,cts_tfs_diff$regulon,paste,collapse=";")
cts_tfs_diff_type_list <- split(names(cts_tfs_diff_type),cts_tfs_diff_type)
#---regulon level
cts_tf_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
fwrite(cts_tf_regulons,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res_aucell.tsv",sep="\t")

cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01 & CSS >0.3 & exp_percent>0.2)
fwrite(cts_tf_regulons_diff,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_diff_res.tsv",sep="\t")
cts_tf_regulons_diff_stat <- table(cts_tf_regulons_diff$regulon,cts_tf_regulons_diff$stage)
attributes(cts_tf_regulons_diff_stat)$class <- "matrix"
data_set <- as.data.frame(cts_tf_regulons_diff_stat)
data_set <- data_set[,unique(colData(atfr)$stage)]

p2 <- upset(data_set, nsets = 6, nintersects = 40, mb.ratio = c(0.5, 0.5),
      order.by = c("degree","freq"), decreasing = c(F,TRUE),point.size=5,
      line.size=2.5)
cts_tf_regulons_diff_type <- tapply(cts_tf_regulons_diff$stage,cts_tf_regulons_diff$regulon,paste,collapse=";")
cts_tf_regulons_diff_type_list <- split(names(cts_tf_regulons_diff_type),cts_tf_regulons_diff_type)

overlap_stages <- intersect(names(cts_tfs_diff_type_list),names(cts_tf_regulons_diff_type_list))
overlap_stages_overlap_targets <- purrr::map2(.x = cts_tfs_diff_type_list[overlap_stages],
                                  .y = cts_tf_regulons_diff_type_list[overlap_stages],
                                  .f = intersect)
sapply(overlap_stages_overlap_targets,length)

overlap_stages_regulon_only_targets <- purrr::map2(.x = cts_tfs_diff_type_list[overlap_stages],
                                  .y = cts_tf_regulons_diff_type_list[overlap_stages],
                                  .f = function(x,y){setdiff(y,x)})

#---candidates plot
p1 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Sox17", by_exprs_values = "logcounts", point_size=0.2,point_alpha=1)
p2 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Sox17", by_exprs_values = "active_score", point_size=0.2,point_alpha=1)
p1+p2

ggsave(filename ="Sox17_tSNE.pdf" ,plot = p1+p2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_candidates",
       width = 110, height =40, dpi = 600, limitsize = FALSE, units ="mm")
#---Irf7
p1 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Irf7", by_exprs_values = "logcounts", point_size=0.2,point_alpha=1)
p2 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Irf7", by_exprs_values = "active_score", point_size=0.2,point_alpha=1)
p1+p2

ggsave(filename ="Irf7_tSNE.pdf" ,plot = p1+p2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_candidates",
       width = 110, height =40, dpi = 600, limitsize = FALSE, units ="mm")

#---Spi1
p1 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Spi1", by_exprs_values = "logcounts", point_size=0.2,point_alpha=1)
p2 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Spi1", by_exprs_values = "active_score", point_size=0.2,point_alpha=1)
p1+p2

ggsave(filename ="Spi1_tSNE.pdf" ,plot = p1+p2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_candidates",
       width = 115, height =40, dpi = 600, limitsize = FALSE, units ="mm")

#---Hlf
p1 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Hlf", by_exprs_values = "logcounts", point_size=0.2,point_alpha=1)
p2 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = "Hlf", by_exprs_values = "active_score", point_size=0.2,point_alpha=1)
p1+p2

ggsave(filename ="Hlf_tSNE.pdf" ,plot = p1+p2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_candidates",
       width = 110, height =40, dpi = 600, limitsize = FALSE, units ="mm")



```


#### 3.2 cell type specific TF regulons but not cts at expression level

```{r}
cts_tfs <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
cts_tfs_diff <- cts_tfs %>% filter(padj<0.01 & CSS >0.3)

cts_tf_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01 & CSS >0.3 & exp_percent>0.2)

get_diff_candidates <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  not_diff_tf <- tf %>% filter(stage %in% stages & padj > 0.3)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  not_diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj > 0.5)
  overlap_tfs <- intersect(diff_tf_regulon$regulon, not_diff_tf$regulon)
  diff_tf_regulon_tab <- diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  all_tf_exp_tab <- not_diff_tf %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_regulon_tab,all_tf_exp_tab,by=c("regulon"="regulon"))
  out_tbl
}

stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
cts_regulons_not_tfs_list <- lapply(stages,get_diff_candidates)
cts_regulons_not_tfs_list <- do.call(rbind,cts_regulons_not_tfs_list)
fwrite(cts_regulons_not_tfs_list,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_regulon_at_activity_level_but_not_expression_level_v2.tsv",sep="\t")

get_diff_candidates2 <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  not_diff_tf <- tf %>% filter(stage %in% stages & padj > 0.3)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  not_diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj > 0.5)
  overlap_tfs <- intersect(diff_tf_regulon$regulon, diff_tf$regulon)
  diff_tf_regulon_tab <- diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  all_tf_exp_tab <- diff_tf %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_regulon_tab,all_tf_exp_tab,by=c("regulon"="regulon"))
  out_tbl
}
stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
cts_regulons_and_tfs_list <- lapply(stages,get_diff_candidates2)
cts_regulons_and_tfs_list <- do.call(rbind,cts_regulons_and_tfs_list)
fwrite(cts_regulons_and_tfs_list,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_regulon_at_activity_level_and_expression_level_v2.tsv",sep="\t")


exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")
assay(act_sce,i = "logcounts") <- assay(exp_sce,"logcounts")[row.names(act_sce),]

i <- "Hoxa1"
i <- "Klf3"
i <- "Hes1"
i <- "Gabpa"
i <- "Kmt2a"
i <- "Srebf2"
i <- "Srebf1"
i <- "Eea1"
i <- "Foxo1"
i <- "Zbtb7a"
i <- "Rela"
p1 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = i, by_exprs_values = "active_score", point_size=1,point_alpha=1)
p2 <- plotReducedDim(object = act_sce,dimred = "TSNE", colour_by = i, by_exprs_values = "logcounts", point_size=1,point_alpha=1)
p1+p2

i <- c("Zbtb7a","Ncoa2","Elf2","Elk4","Gtf2b")
i <- cts_regulons_not_tfs_list %>% filter(stage.x =="HEC") %>% pull(regulon)


i <- "Sp1"
v1 <- plotExpression(object = act_sce,x = "stage",exprs_values = "active_score",colour_by="stage",features = i, point_size=0.5, point_alpha=1,ncol = round(sqrt(length(i))),scales="free_y")
v2 <- plotExpression(object = act_sce,x = "stage",exprs_values = "logcounts",colour_by="stage",features = i, point_size=0.5, point_alpha=1,ncol = round(sqrt(length(i))),scales="free_y")
v2+v1
ggsave(filename ="Sp1_expression_violin.pdf" ,plot = v2+v1, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/figure/",
       width = 200, height =75, dpi = 600, limitsize = FALSE, units ="mm")


i <- "Nfxl1"
i <- "Ncoa2"
i <- "Runx1"
i <- cts_regulons_not_tfs_list %>% filter(stage.x =="FL_HSC") %>% pull(regulon)


v1 <- plotExpression(object = act_sce,x = "stage",exprs_values = "active_score",colour_by="stage",features = i, point_size=0.1, point_alpha=1,ncol = round(sqrt(length(i))),scales="free_y")
ggsave(filename ="FL_diff_regulon_not_gene_activate_score.pdf" ,plot = v1, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/candidates",
       width = 200, height =150, dpi = 600, limitsize = FALSE, units ="mm")
v2 <- plotExpression(object = act_sce,x = "stage",exprs_values = "logcounts",colour_by="stage",features = i,point_size=0.1, point_alpha=1,ncol = round(sqrt(length(i))),scales="free_y")
ggsave(filename ="FL_diff_regulon_not_gene_expression.pdf" ,plot = v2, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/candidates",
       width = 200, height =150, dpi = 600, limitsize = FALSE, units ="mm")
```


#### 3.3 cell type specific TF regulons but not cts at expression level comparison

```{r}
cts_tfs <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
cts_tfs_diff <- cts_tfs %>% filter(padj<0.01 & CSS >0.3)

cts_tf_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01 & CSS >0.3 & exp_percent>0.3)

#---expression and regulon level
get_overlap <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  overlap_tfs <- intersect(diff_tf$regulon, diff_tf_regulon$regulon)
  diff_tf_tab <- diff_tf %>% filter(regulon %in% overlap_tfs)
  diff_tf_regulon_tab <- diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_tab,diff_tf_regulon_tab,by=c("regulon"="regulon"))
  out_tbl
}
stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
cts_tf_regulons_list <- lapply(stages, get_overlap)
names(cts_tf_regulons_list) <- stages
cts_tf_regulons_tbl <- do.call(rbind,cts_tf_regulons_list)
length(table(cts_tf_regulons_tbl$regulon))

#---regulon level only
get_diff_regulons <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  not_diff_tf <- tf %>% filter(stage %in% stages & padj > 0.5)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  not_diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj > 0.5)
  overlap_tfs <- intersect(diff_tf_regulon$regulon, not_diff_tf$regulon)
  diff_tf_regulon_tab <- diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  all_tf_exp_tab <- not_diff_tf %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_regulon_tab,all_tf_exp_tab,by=c("regulon"="regulon"))
  out_tbl
}
cts_regulons_not_tfs_list <- lapply(stages,get_diff_regulons)
names(cts_regulons_not_tfs_list) <- stages
cts_regulons_not_tfs_tbl <- do.call(rbind,cts_regulons_not_tfs_list)
length(table(cts_regulons_not_tfs_tbl$regulon))

#---expression level only
get_diff_tfs <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  not_diff_tf <- tf %>% filter(stage %in% stages & padj > 0.5)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  not_diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj > 0.5)
  overlap_tfs <- intersect(diff_tf$regulon, not_diff_tf_regulon$regulon)
  diff_tf_tab <- diff_tf %>% filter(regulon %in% overlap_tfs)
  all_tf_exp_tab <- not_diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_tab, all_tf_exp_tab,by=c("regulon"="regulon"))
  out_tbl
}
cts_tfs_not_regulons_list <- lapply(stages,get_diff_tfs)
names(cts_tfs_not_regulons_list) <- stages
cts_tfs_not_regulons_tbl <- do.call(rbind,cts_tfs_not_regulons_list)
length(table(cts_tfs_not_regulons_tbl$regulon))

#---sig regulon each stage 
cts_tf_regulons_num <- tapply(cts_tf_regulons_diff$regulon,cts_tf_regulons_diff$stage,length)
cts_tfs_num <- tapply(cts_tfs_diff$regulon,cts_tfs_diff$stage,length)
cts_tfs_regulons_overlap_num <- sapply(cts_tf_regulons_list,nrow)
cts_tfs_regulons_only_num <- sapply(cts_regulons_not_tfs_list,nrow)
cts_tfs_only_num <- sapply(cts_tfs_not_regulons_list,nrow)

regulons_frm <- data.frame(stage = stages,number = cts_tf_regulons_num[stages],type="regulons")
tfs_frm <- data.frame(stage = stages,number = cts_tfs_num[stages],type="TFs")
plot_frm <- rbind(regulons_frm,tfs_frm)
plot_frm$stage <- factor(plot_frm$stage,levels=stages)
plot_frm$type <- factor(plot_frm$type,levels=c("TFs","regulons"))

library(ggplot2)
p <- ggplot(plot_frm,aes(x=stage,y=number,fill=type))+
  geom_bar(stat = "identity",position = "dodge")
p

pie(cts_tfs_regulons_overlap_num,clockwise = T)
pie(cts_tfs_regulons_only_num,clockwise = T)
pie(cts_tfs_only_num,clockwise = T)

```

#### 3.4 CTS regulons pathways

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
all_regulons <- as.list(Regulon(atfr))
score_cst_regulons <- all_regulons

reactome_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/scATFR/inst/extdata/Reactome_pathway.rds")
mouse_reactome <- reactome_pathway$`Mus musculus`

score_cst_regulons_reactome <- calcuModuleSimilar(tfr_list = score_cst_regulons, pathway_list = mouse_reactome, adjust = "BH" ,ncores=6)
saveRDS(score_cst_regulons_reactome,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/pathways/all_regulons_reactome_pathway.rds")
score_cst_regulons_reactome_sig <- score_cst_regulons_reactome %>% filter(padj<0.01)
#score_cst_regulons_reactome_sig %>% filter(regulon %in% "Spi1")

#---both level
cts_tf_regulons_tbl_reactome <- inner_join(cts_tf_regulons_tbl, score_cst_regulons_reactome,by=c("regulon"="regulon"))
cts_tf_regulons_tbl_reactome_diff <- cts_tf_regulons_tbl_reactome %>% filter(firstInGroup==1)

ststs <- table(cts_tf_regulons_tbl_reactome$pathway, cts_tf_regulons_tbl_reactome$stage.x)
attributes(ststs)$class <-"matrix"
ststs <- ststs[rowSums(ststs)>=2,]

library(pheatmap)
p <- pheatmap(ststs[,stages],show_rownames = F,scale="row",cluster_cols = F)


#---regulon level
cts_regulons_not_tf_tbl_reactome <- inner_join(cts_regulons_not_tfs_tbl, score_cst_regulons_reactome_sig, by=c("regulon"="regulon"))
ststs <- table(cts_regulons_not_tf_tbl_reactome$pathway, cts_regulons_not_tf_tbl_reactome$stage.x)
attributes(ststs)$class <-"matrix"
ststs <- ststs[rowSums(ststs)>=2,]
anno_row <- as.data.frame(rowSums(ststs))
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
p <- pheatmap(ststs[,stages], show_rownames = T, scale="row", cluster_cols = F, fontsize = 5, annotation_row = anno_row, color = pal, cutree_rows = 4)

cts_regulons_not_tf_tbl_reactome %>% filter(pathway %in% 'Transcriptional regulation by RUNX1')

cts_regulons_not_tf_tbl_reactome %>% filter(pathway %in% 'RUNX1 regulates transcription of genes involved in differentiation of HSCs')

cts_regulons_not_tf_tbl_reactome %>% filter(pathway %in% c('Cell Cycle','Cell Cycle, Mitotic'))

cts_regulons_not_tf_tbl_reactome %>% filter(pathway %in% c('MAPK family signaling cascades'))

cts_regulons_not_tf_tbl_reactome %>% filter(pathway %in% c('Programmed Cell Death'))

distinc_reactome <- setdiff(cts_tf_regulons_tbl_reactome$pathway,cts_regulons_not_tf_tbl_reactome$pathway)

```





```{r}
out_list <- list(both_level = cts_tf_regulons_tbl, 
                 regulon_only = cts_regulons_not_tfs_tbl,
                 tf_only = cts_tfs_not_regulons_tbl)
saveRDS(out_list,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/CST_TFs_at_three_types.rds")

cts_tfs_order <- cts_tfs[order(cts_tfs$stage,cts_tfs$regulon),]
cts_tfs_order$CSS <- (cts_tfs_order$CSS-min(cts_tfs_order$CSS))/(max(cts_tfs_order$CSS)-min(cts_tfs_order$CSS))
cts_tf_regulons_order <- cts_tf_regulons[order(cts_tf_regulons$stage,cts_tf_regulons$regulon),]
cts_tf_regulons_order$CSS <- (cts_tf_regulons_order$CSS-min(cts_tf_regulons_order$CSS))/(max(cts_tf_regulons_order$CSS)-min(cts_tf_regulons_order$CSS))

cts_tfs_ratio <- (1+cts_tf_regulons_order$CSS)/(cts_tfs_order$CSS+1)
cts_tfs_ratio_tab <- cts_tf_regulons_order[,1:2]
cts_tfs_ratio_tab$ratio <- cts_tfs_ratio
cts_tfs_ratio_tab_dcat <- dcast(cts_tfs_ratio_tab,formula = regulon ~stage)
write.table(cts_tfs_ratio_tab_dcat,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/CSS_ratio_between_regulon_and_expression.tsv",sep="\t",quote=F,row.names=F)
stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
cts_tfs_ratio_tab_dcat <- cts_tfs_ratio_tab_dcat[,c("regulon",stages)]
row.names(cts_tfs_ratio_tab_dcat) <- cts_tfs_ratio_tab_dcat$regulon
cts_tf_regulons %>% filter(regulon %in% "2700081O15Rik",)

get_diff_regulons <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  not_diff_tf <- tf %>% filter(stage %in% stages & padj > 0.5)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  not_diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj > 0.5)
  overlap_tfs <- intersect(diff_tf_regulon$regulon, not_diff_tf$regulon)
  diff_tf_regulon_tab <- diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  all_tf_exp_tab <- not_diff_tf %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_regulon_tab,all_tf_exp_tab,by=c("regulon"="regulon"))
  out_tbl
}

get_diff_tfs <- function(stages,tf=cts_tfs, regulon=cts_tf_regulons){
  diff_tf <- tf %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3)
  not_diff_tf <- tf %>% filter(stage %in% stages & padj > 0.5)
  diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj < 0.01 & CSS >0.3 & exp_percent >= 0.3)
  not_diff_tf_regulon <- regulon %>% filter(stage %in% stages & padj > 0.2)
  overlap_tfs <- intersect(diff_tf$regulon, not_diff_tf_regulon$regulon)
  diff_tf_tab <- diff_tf %>% filter(regulon %in% overlap_tfs)
  all_tf_exp_tab <- not_diff_tf_regulon %>% filter(regulon %in% overlap_tfs)
  out_tbl <- inner_join(diff_tf_tab, all_tf_exp_tab,by=c("regulon"="regulon"))
  out_tbl
}

stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
cts_regulons_not_tfs_list <- lapply(stages,get_diff_candidates)
cts_regulons_not_tfs_list <- do.call(rbind,cts_regulons_not_tfs_list)


cts_tfs_not_regulons_list <- lapply(stages,get_diff_tfs)
cts_tfs_not_regulons_list <- do.call(rbind,cts_tfs_not_regulons_list)

cts_tfs_ratio_tab_dcat["Hlf",]
cts_tfs %>% filter(regulon %in% "Hlf")

cts_regulons_not_tfs_list$ratio <- cts_regulons_not_tfs_list$CSS.x/cts_regulons_not_tfs_list$CSS.y



cts_regulons_type <- tapply(cts_regulons_not_tfs_list$stage.x,cts_regulons_not_tfs_list$regulon,function(x){paste(x,collapse = ";")})
type_all <- names(table(cts_regulons_type))
new_type <- c("Adult_HSC","AEC","Adult_HSC","HEC","FL_HSC","Adult_HSC","HEC","T1_pre_HSC","T1_pre_HSC","T2_pre_HSC", "FL_HSC")
names(new_type) <- type_all
cts_regulons_type_new <- new_type[cts_regulons_type]
names(cts_regulons_type_new) <- names(cts_regulons_type)



#---get mean score
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
act_cst_score <- assay(altExp(atfr,"viper"))
col_data <- colData(atfr)
sig_act_cst_score <- act_cst_score[unique(cts_regulons_not_tfs_list$regulon),]
sig_act_cst_score <- t(apply(sig_act_cst_score,1,function(x){(x-min(x))/(max(x)-min(x))}))
sig_act_cst_score_stage <- t(apply(sig_act_cst_score,1,function(x){
  tapply(x,col_data[names(x),"stage"],median)
}))

act_cst_logcounts <- assay(atfr,"logcounts")[row.names(act_cst_score),]



sig_act_cst_score_stage["Foxc1",]
stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
sig_act_cst_score_stage <- cts_tfs_ratio_tab_dcat[unique(cts_regulons_not_tfs_list$regulon),stages]
sig_act_cst_score_stage <- as.data.frame(sig_act_cst_score_stage)
sig_act_cst_score_stage$type <- factor(cts_regulons_type_new[row.names(sig_act_cst_score_stage)],levels=stages)
sig_act_cst_score_stage$mean <- apply(sig_act_cst_score_stage[,stages],1,mean)

plot_frm <- sig_act_cst_score_stage
rv <- getRadviz(input_frm = plot_frm,anchors =stages,optim_anchor=F,text_size=6)
p <- do_RadvizEnrichPlot(rv,color_by="HEC",alpha = 0.8,label_size = 1)+
  scale_color_manual(values=c("#039be5","#ffc107","#f44336","#4caf50","#9c27b0","#3f51b5"))
p
ggsave(filename ="cts_regulon_not_gene_expression.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/figure",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")

```


#### 3.4 cell type specific TF regulons at T1 & T2


```{r}
cts_tfs <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tfs_all_res.rds")
cts_tfs_diff <- cts_tfs %>% filter(padj<0.01 & CSS >0.3)

cts_tf_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_all_res.rds")
cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01 & CSS >0.3 & exp_percent>0.3)

atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
col_data <- colData(atfr)
regulon_activaty <- assay(altExp(atfr,"viper"))
regulon_activaty_cv <- t(apply(regulon_activaty,1,function(x){
    cv <- tapply(x,col_data[names(x),"stage"],function(y){
    y <- (y-min(y))/(max(y)-min(y))
  (sd(y)+0.001)/(mean(y)+0.001)
  })
  ecv <- ecdf(cv)(cv)
  names(ecv) <- names(cv)
  ecv 
  }))
logcounts_exp <- assay(atfr,"logcounts")[row.names(regulon_activaty),]
logcounts_exp_cv <- t(apply(logcounts_exp,1,function(x){
  cv <- tapply(x,col_data[names(x),"stage"],function(y){
    y <- (y-min(y))/(max(y)-min(y))
  (sd(y)+0.001)/(mean(y)+0.001)
  })
  ecv <- ecdf(cv)(cv)
  names(ecv) <- names(cv)
  ecv
  }))
cts_tfs$regulon_cv <- apply(cts_tfs,1,function(x,regulon_activaty_cv){
    regulon_activaty_cv[x[[2]],x[[1]]]
},regulon_activaty_cv = regulon_activaty_cv)
cts_tfs$exp_cv <- apply(cts_tfs,1,function(x,logcounts_exp_cv){
  logcounts_exp_cv[x[[2]],x[[1]]]
},logcounts_exp_cv = logcounts_exp_cv)

cts_tf_regulons$regulon_cv <- apply(cts_tf_regulons,1,function(x,regulon_activaty_cv){
  regulon_activaty_cv[x[[2]],x[[1]]]
},regulon_activaty_cv = regulon_activaty_cv)
cts_tf_regulons$exp_cv <- apply(cts_tf_regulons,1,function(x,logcounts_exp_cv){
  logcounts_exp_cv[x[[2]],x[[1]]]
},logcounts_exp_cv = logcounts_exp_cv)

cts_tfs_diff <- cts_tfs %>% filter(padj<0.01 & CSS >0.3)
cts_tf_regulons_diff <- cts_tf_regulons %>% filter(padj<0.01 & CSS >0.3 & exp_percent>0.3)




cts_list <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/CST_res/CST_TFs_at_three_types.rds")
cts_list_tbl <- data.table::rbindlist(cts_list_tbl,fill =T,idcol = "type")
cts_list_tbl$regulon_cv <- apply(cts_list_tbl,1,function(x,regulon_activaty_cv){
  regulon_activaty_cv[x[[3]],x[[2]]]
},regulon_activaty_cv = regulon_activaty_cv)
cts_list_tbl$exp_cv <- apply(cts_list_tbl,1,function(x,logcounts_exp_cv){
  logcounts_exp_cv[x[['regulon']],x[['stage.x']]]
},logcounts_exp_cv = logcounts_exp_cv)



plot_frm <- cts_tfs[,c('regulon_cv', 'exp_cv')]
plot_frm <- cts_tfs_diff[,c('regulon_cv', 'exp_cv')]

plot_frm <- cts_list_tbl[,c('regulon_cv', 'exp_cv','type')]


library(ggplot2)
p <- ggplot(data = plot_frm)+
  stat_ecdf(mapping = aes(x=exp_cv),color="blue")+
  stat_ecdf(mapping = aes(x=regulon_cv),color="red")+
  scale_x_continuous(limits=c(0,3))+
  facet_wrap(.~type)
p
ks.test(plot_frm$logcounts_exp_cv,plot_frm$regulon_activaty_cv)

```





