---
title: "R Notebook"
output: html_notebook
---
#### 1.1 load data and create object

```{r}
library(scATFR)
library(PIDC)
exp_data <- readRDS(system.file("./inst/extdata/mouse_HSC_formation_expression.rds",package = "scATFR"))
col_metadata <- readRDS(system.file("./inst/extdata/mouse_HSC_formation_colData.rds",package = "scATFR"))
row.names(col_metadata) <- col_metadata$sample
atfr <- scATFR(exp_data = exp_data,col_data = col_metadata)

```

#### 1.2 Inferring gene regulatory network

```{r}
atfr <- inferGRNs(x = atfr, method="pidc",ncores=6)
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/GRNs/atfr_pidc_grn.rds")
GRNs(atfr)
TF_gene <- intersect(TF_Symbols$mouse,row.names(atfr))
PIDC_nets <- as.matrix(GRN(atfr,"pidc")) %>% reshape2::melt(.) %>% dplyr::filter(Var1 %in% TF_gene)
colnames(PIDC_nets) <- c("tf","target","weight")
PIDC_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/GRNs/pidc_grn_regulons.rds") %>%
  filter(regulator %in% TF_gene)
#---load regulons
blood_network <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/final_network/mouse/blood_merged_filter_strict_network.txt.gz",sep="\t",header=T) %>% filter(tf %in% TF_gene)
#blood_network <- filterNets(df = blood_network, minSize = 50, maxSize = 2000, maxInter = 10,ncores = 6)
#dim(blood_network[blood_network$tf=="Batf2",])
#dim(blood_network[blood_network$tf=="Nkx2-1",])
#dim(blood_network[blood_network$tf=="Runx1",])

regulons <- integraNets(grnNets = list(PIDC_regulons, blood_network), norm_weight = T,
                        max_target_num=2000, min_target_num = 100, join_type = "full", 
                        return_sig = T, maxInter = 5, ncores = 2)
interged_regulon_list <- dfToList(regulons)
atfr <- filterRegulons(x = atfr, gene_list=interged_regulon_list, maxSize=500, minSize=5)
sapply(Regulon(atfr),length)
atfr <- regulonActivity(x = atfr, method="viper")
active_mat <- assay(altExp(atfr,"viper"))
write.table(active_mat,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/TF_active_mat.txt",sep="\t",quote=F)
saveRDS(atfr,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
```
#### 2.1 quality control

```{r}
exp_data <- readRDS(system.file("./inst/extdata/mouse_HSC_formation_expression.rds",package = "scATFR"))
col_metadata <- readRDS(system.file("./inst/extdata/mouse_HSC_formation_colData.rds",package = "scATFR"))
row.names(col_metadata) <- col_metadata$sample
gene_number <- apply(exp_data,2,function(x){length(x[x>=1])})
col_metadata$gene_number <- gene_number[row.names(col_metadata)]
col_metadata$stage <- factor(col_metadata$stage,levels=unique(col_metadata$stage))
stage2color <- read.table("D:/HSC_Autophagy/2.HSC_formation/sample_infor/stage2color.txt",sep="\t",header=T,comment.char = "")
row.names(stage2color) <- stage2color$stage

text_size = 8
line_size = 0.5/1.07
library(ggplot2)
#------set plot theme
mytheme = theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA,color=NA),
strip.background = element_rect(fill=NA,color=NA),
strip.text = element_text(size=text_size,color="black"),
panel.grid.major = element_line(size=line_size/2, linetype = 8, color = "lightgray"),
panel.grid.minor = element_line(size=NA, color = NA),
axis.ticks = element_line(size=line_size,color="black"),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45, hjust = 1, vjust =1,color="black"), 
axis.text.y= element_text(size=text_size,color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

plot_frm <- col_metadata

p <- ggplot(plot_frm,mapping = aes(x=stage,y=gene_number, color=stage))+
geom_violin(position = "dodge", na.rm = TRUE, scale = "width",width = 0.9)+
geom_boxplot(width=0.25,fill=NA, size=0.2,outlier.colour=NA, alpha = 0.8)+
geom_jitter(width=0.2,size=0.8,shape=16,color="black", alpha = 0.8)+
  scale_color_manual(values = stage2color$figurecolor)+
labs(x="",y='number of expressed genes in each single cells')+
mytheme+
  scale_y_continuous(limits = c(0,10000))

p

saveRDS(p,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/quality_control/gene_number_violin.rds")

```

#### 2.2 dimensionality reduction using expression data and activate score

```{r}
library(scater)
library(scran)
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
tf_gene <- intersect(TF_Symbols$mouse,row.names(atfr))
exp_sce <- SingleCellExperiment::SingleCellExperiment(list(counts=as.matrix(assay(atfr)[tf_gene,])))
exp_sce <- scater::logNormCounts(exp_sce)
colData(exp_sce) <- colData(atfr)
colData(exp_sce)$stage <- reLevel(vect = colData(exp_sce)$stage)
exp_sce <- scater::runPCA(x=exp_sce,subset_row=tf_gene) 
plotReducedDim(exp_sce, dimred="PCA",colour_by = "stage")
exp_sce <- scater::runTSNE(x = exp_sce,subset_row=tf_gene)
plotReducedDim(exp_sce, dimred="TSNE",colour_by = "stage")
exp_sce  <- scater::runUMAP(x = exp_sce, subset_row=tf_gene)
plotReducedDim(exp_sce , dimred="UMAP",colour_by = "stage")
saveRDS(exp_sce,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.1_TF_exp_sce.rds")

#--- active score
tf_gene <- intersect(TF_Symbols$mouse,row.names(atfr))
act_sce <- SingleCellExperiment::SingleCellExperiment(list(active_score=assay(altExp(atfr,"viper"))))
colData(act_sce) <- colData(atfr)
colData(act_sce)$stage <- reLevel(vect = colData(act_sce)$stage)
act_sce <- scater::runPCA(x=act_sce,subset_row=tf_gene,exprs_values="active_score") 
plotReducedDim(act_sce, dimred="PCA",colour_by = "stage")
act_sce <- scater::runTSNE(x = act_sce,subset_row=tf_gene,exprs_values="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "stage")
act_sce <- scater::runUMAP(x = act_sce, subset_row=tf_gene,exprs_values="active_score")
plotReducedDim(act_sce, dimred="UMAP",colour_by = "stage")
saveRDS(act_sce,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.2_TF_act_sce.rds")

```

#### 2.3 cell type specific using TF activities

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
col_data <- as.data.frame(colData(atfr))
col_data$stage <- as.character(col_data$stage)
col_data$stage <- ifelse(col_data$stage %in% c("E12","E14"),"FL_HSC",col_data$stage)
colData(atfr)$stage <- col_data$stage
cst_regulon <- findAllCTSRegulons(x = atfr, use_assay="viper", clust_name ="stage",
                              ncores=6, nperm=10000,fdr=1,css=0,expression_per = 0.0)
cst_mat <- diffRegulon(cst_regulon)
cst_mat_sig <- cst_mat %>% filter(padj<0.01 & CSS >=0.35 & exp_percent>0.1)

#---expression level
TF_exp_mat <- as.matrix(assay(atfr,"counts"))[row.names(assay(altExp(atfr,"viper"))),]
exp_cst <- findAllCTSRegulons(x = TF_exp_mat, col_data = col_data, scale=T,
                          clust_name ="stage",ncores=6, nperm=10000,fdr=1, css=0)
exp_cst_sig <- exp_cst %>% filter(padj<0.01 & CSS >=0.35)
saveRDS(list(exp_cst=exp_cst,score_cst=cst_mat),"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_all_test_res.rds")
#readr::write_delim(exp_cst,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/diff_res/7_stage_expression_level_cst_all_res.txt",delim = "\t")
#readr::write_delim(cst_regulon,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/diff_res/7_stage_regulon_level_cst_all_res.txt",delim = "\t")
saveRDS(list(exp_cst=exp_cst_sig,score_cst=cst_mat_sig),"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_fdr0.01_CSS_0.35.rds")

```

#### 2.4 cell type specific TF vs regulon histogram

```{r}
#----differential TF number comparison 
library(ggplot2)
p <- ggplot()+
  geom_histogram(data = cst_mat,mapping = aes(x=CSS,y=..count..),bins = 50,fill="red",color="black",alpha=0.6)+
  geom_histogram(data = exp_cst,mapping = aes(x=CSS,y=..count..),bins = 50,fill="blue",color="black",alpha=0.6)+
  scale_x_continuous(breaks=seq(0.0,0.6,0.2))
p
```
#### 2.5 cell type specific regulon top 500 Radviz

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_list <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_all_test_res.rds")
cst_mat_sig <- cst_list$score_cst %>% filter(padj<0.01 & CSS>0.35 & exp_percent >0.1)
act_cst_score <- assay(altExp(atfr,"viper"))
sig_act_cst_score <- act_cst_score[unique(cst_mat_sig$regulon),]
sig_act_cst_score_stage <- t(apply(sig_act_cst_score,1,function(x){
  tapply(x,col_data[names(x),"stage"],median)
}))
regulon_row <- tapply(cst_mat_sig$stage,cst_mat_sig$regulon,function(x){paste(x,collapse = "_")})
#write.table(as.data.frame(regulon_row),"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_regulon_stage_type.txt",quote=F,sep="\t")

stages <- c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
sig_act_cst_score_stage <- sig_act_cst_score_stage[,stages]
sig_act_cst_score_stage <- as.data.frame(sig_act_cst_score_stage)
regulon_row <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_regulon_stage_type.txt",sep="\t",header=T,row.names=1)
sig_act_cst_score_stage$type <- regulon_row[row.names(sig_act_cst_score_stage),"cluster"]
plot_frm <- sig_act_cst_score_stage
rv <- getRadviz(input_frm = plot_frm,anchors =stages,optim_anchor=F,text_size=6)
p <- do_RadvizEnrichPlot(rv,color_by="type",alpha = 0.7,label_size = 3)+
  scale_color_manual(values=c("#729ece","#ff9e4a","#67bf5c","#ed665d","#ad8bc9","#7b1fa2","#f4511e"))
p

```

#### 2.6 cell type specific regulon function

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_test <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_fdr0.01_CSS_0.35.rds")
exp_cst_sig <- cst_test$exp_cst
score_cst_sig <- cst_test$score_cst

score_cst_regulons <- as.list(Regulon(atfr)[unique(score_cst_sig$regulon)])
score_cst_sig_EC <- score_cst_sig %>% filter(stage=="EC")
score_cst_sig_T1 <- score_cst_sig %>% filter(stage=="T1_pre_HSC")
score_cst_sig_FL <- score_cst_sig %>% filter(stage=="FL_HSC")
score_cst_sig_Adult <- score_cst_sig %>% filter(stage=="Adult_HSC")
length(table(score_cst_sig$regulon))
exp_cst_sig_EC <- exp_cst_sig %>% filter(stage=="EC")
exp_cst_sig_T1 <- exp_cst_sig %>% filter(stage=="T1_pre_HSC")
exp_cst_sig_FL <- exp_cst_sig %>% filter(stage=="FL_HSC")
exp_cst_sig_Adult <- exp_cst_sig %>% filter(stage=="Adult_HSC")
length(table(exp_cst_sig$regulon))
overlap_cst_regulon <- intersect(names(table(exp_cst_sig$regulon)),names(table(score_cst_sig$regulon)))

reactome_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/scATFR/inst/extdata/Reactome_pathway.rds")
mouse_reactome <- reactome_pathway$`Mus musculus`
score_cst_regulons_reactome <- calcuModuleSimilar(tfr_list = score_cst_regulons, pathway_list = mouse_reactome, adjust = "BH" ,ncores=6)
score_cst_regulons_reactome_sig <- score_cst_regulons_reactome %>% filter(firstInGroup==1)

merge_enrich_tab <- inner_join(score_cst_sig,score_cst_regulons_reactome_sig,by="regulon")
merge_enrich_tab <- unique(merge_enrich_tab[,c("stage","regulon","pathway")])
regulon_level <- names(sort(table(merge_enrich_tab$regulon)))
pathway_level <- names(sort(table(merge_enrich_tab$pathway)))
stage_level <- c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
merge_enrich_tab$stage <- factor(merge_enrich_tab$stage,levels=rev(stage_level))
merge_enrich_tab$regulon <- factor(merge_enrich_tab$regulon,levels=regulon_level)
merge_enrich_tab$pathway <- factor(merge_enrich_tab$pathway,levels=pathway_level)

library("easyalluvial")
library("parcats")
data <- merge_enrich_tab
p = alluvial_wide(data, max_variables = 3, fill_by = "first_variable",
                  colorful_fill_variable_stratum=F,col_vector_value=NA)
parcats(p, marginal_histograms = FALSE, data_input = data,height=1200,width = 400)

heatmap_stat <- table(merge_enrich_tab$pathway,merge_enrich_tab$stage)
attributes(heatmap_stat)$calss <- "matrix"
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
heatmap_stat <- heatmap_stat[rowSums(heatmap_stat)>=3,stage_level]
anno_row <- as.data.frame(rowSums(heatmap_stat))

pheatmap(heatmap_stat,show_rownames = T,scale="row",cluster_cols = F,fontsize = 4,treeheight_row = F,
         clustering_method = "ward.D2",color = pal, display_numbers = F,annotation_row = anno_row)
merge_enrich_tab %>% filter(pathway %in% 'Transcriptional regulation by RUNX1')
merge_enrich_tab %>% filter(pathway %in% 'mRNA Splicing')
merge_enrich_tab %>% filter(pathway %in% 'Transcriptional Regulation by TP53')
merge_enrich_tab %>% filter(pathway %in% 'Nuclear Receptor transcription pathway')
merge_enrich_tab %>% filter(pathway %in% 'Signaling by MET')
merge_enrich_tab %>% filter(pathway %in% 'Processing of Capped Intron-Containing Pre-mRNA')
merge_enrich_tab %>% filter(pathway %in% 'Immune System')
merge_enrich_tab %>% filter(pathway %in% 'Signaling by the B Cell Receptor (BCR)')
merge_enrich_tab %>% filter(pathway %in% 'Platelet homeostasis')
merge_enrich_tab %>% filter(pathway %in% 'Signaling by MET')
merge_enrich_tab %>% filter(pathway %in% 'MAPK family signaling cascades')
merge_enrich_tab %>% filter(pathway %in% 'Cytokine Signaling in Immune system')
merge_enrich_tab %>% filter(pathway %in% 'DNA Repair')
merge_enrich_tab %>% filter(pathway %in% 'Gene expression (Transcription)')
merge_enrich_tab %>% filter(pathway %in% 'Cell Cycle')
merge_enrich_tab %>% filter(pathway %in% 'RNA Polymerase II Transcription')
merge_enrich_tab %>% filter(pathway %in% 'Deubiquitination')
```

#### 2.7 cell type specific TF function

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_test <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_fdr0.01_CSS_0.35.rds")
exp_cst_sig <- cst_test$exp_cst
score_cst_sig <- cst_test$score_cst
score_cst_regulons <- as.list(Regulon(atfr)[unique(score_cst_sig$regulon)])
score_cst_sig_EC <- score_cst_sig %>% filter(stage=="EC")
score_cst_sig_T1 <- score_cst_sig %>% filter(stage=="T1_pre_HSC")
score_cst_sig_T2 <- score_cst_sig %>% filter(stage=="T2_pre_HSC")
score_cst_sig_FL <- score_cst_sig %>% filter(stage=="FL_HSC")
score_cst_sig_Adult <- score_cst_sig %>% filter(stage=="Adult_HSC")

exp_cst_sig_EC <- exp_cst_sig %>% filter(stage=="EC")
exp_cst_sig_T1 <- exp_cst_sig %>% filter(stage=="T1_pre_HSC")
exp_cst_sig_T2 <- exp_cst_sig %>% filter(stage=="T2_pre_HSC")
exp_cst_sig_FL <- exp_cst_sig %>% filter(stage=="FL_HSC")
exp_cst_sig_Adult <- exp_cst_sig %>% filter(stage=="Adult_HSC")

exp_cst_list <- list(EC=exp_cst_sig_EC$regulon,
                        T1_T2=unique(c(exp_cst_sig_T1$regulon,exp_cst_sig_T2$regulon)),
                        FL=exp_cst_sig_FL$regulon,Adult=exp_cst_sig_Adult$regulon)

reactome_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/scATFR/inst/extdata/Reactome_pathway.rds")
mouse_reactome <- reactome_pathway$`Mus musculus`
exp_cst_tf_reactome <- calcuModuleSimilar(tfr_list = exp_cst_list, pathway_list = mouse_reactome, adjust = "BH" ,ncores=6, min_overlap = 1)
select_terms <- c('Deactivation of the beta-catenin transactivating complex',
                  'Transcriptional regulation by RUNX3',
                  'Transcriptional activation of mitochondrial biogenesis',
                  'DNA Damage/Telomere Stress Induced Senescence',
                  'G0 and Early G1',
                  'MyD88-independent TLR4 cascade',
                  'Nuclear Receptor transcription pathway')
exp_cst_tf_reactome_sig <- exp_cst_tf_reactome %>% filter(pathway %in% select_terms)
exp_cst_tf_reactome_sig$regulon <- factor(exp_cst_tf_reactome_sig$regulon,levels=c("EC","T1_T2","FL","Adult"))
plot_frm <- exp_cst_tf_reactome_sig
library(ggplot2)
p <- ggplot(plot_frm,aes(x = pathway,y = regulon, size=similarity, color=padj))+
  geom_point()+
  scale_color_gradient(low = '#1565c0',high = '#c62828')+
  coord_flip()
p

```


#### 2.8 cell type specific regulon vs expression level

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_test <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_all_test_res.rds")
exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.2_TF_act_sce.rds")
assay(act_sce,"logcounts") <- assay(exp_sce,"logcounts")[row.names(assay(act_sce,"active_score")),]
exp_cst <- cst_test$exp_cst 
score_cst <- cst_test$score_cst 

score_cst_sig <- score_cst %>% filter(padj<0.01 & CSS>=0.35 & exp_percent>=0.1)
score_cst_regulons <- as.list(Regulon(atfr)[unique(score_cst_sig$regulon)])
score_cst_sig_EC <- score_cst_sig %>% filter(stage=="EC")
score_cst_sig_T1 <- score_cst_sig %>% filter(stage=="T1_pre_HSC")
score_cst_sig_T2 <- score_cst_sig %>% filter(stage=="T2_pre_HSC")
score_cst_sig_FL <- score_cst_sig %>% filter(stage=="FL_HSC")
score_cst_sig_Adult <- score_cst_sig %>% filter(stage=="Adult_HSC")

exp_cst_sig <- exp_cst %>% filter(padj<0.01 & CSS>=0.35)
exp_cst_not_sig <- exp_cst[exp_cst$padj>0.5,]
exp_cst_sig_EC <- exp_cst_sig %>% filter(stage=="EC")
exp_cst_sig_T1 <- exp_cst_sig %>% filter(stage=="T1_pre_HSC")
exp_cst_sig_T2 <- exp_cst_sig %>% filter(stage=="T2_pre_HSC")
exp_cst_sig_FL <- exp_cst_sig %>% filter(stage=="FL_HSC")
exp_cst_sig_Adult <- exp_cst_sig %>% filter(stage=="Adult_HSC")

#---T1 T2 specific
T1_T2_regulon_spec <- unique(c(score_cst_sig_T1$regulon,score_cst_sig_T2$regulon))
T1_T2_exp_spec <- unique(c(exp_cst_sig_T1$regulon,exp_cst_sig_T2$regulon))
T1_T2_exp_regulon_overlap <- intersect(T1_T2_regulon_spec,T1_T2_exp_spec)
T1_T2_regulon_setdiff <- setdiff(T1_T2_regulon_spec,T1_T2_exp_spec)
T1_T2_samples <- row.names(colData(act_sce)[colData(act_sce)$stage %in% c("T1_pre_HSC","T2_pre_HSC"),])
logcounts_exp <- assay(act_sce,"logcounts")[T1_T2_regulon_setdiff,T1_T2_samples]
logcounts_exp_cv <- apply(logcounts_exp,1,function(x){
  x <- (x-min(x))/(max(x)-min(x))
  sd(x)/mean(x)})
regulon_activaty <- assay(act_sce,"active_score")[T1_T2_regulon_setdiff,T1_T2_samples]
regulon_activaty_cv <- apply(regulon_activaty,1,function(x){
    x <- (x-min(x))/(max(x)-min(x))
    sd(x)/mean(x)})
plot_frm <- data.frame(logcounts_exp_cv, regulon_activaty_cv)
library(ggplot2)
p <- ggplot(data = plot_frm)+
  stat_ecdf(mapping = aes(x=logcounts_exp_cv),color="blue")+
  stat_ecdf(mapping = aes(x=regulon_activaty_cv),color="red")+
  scale_x_continuous(limits=c(0,3))
p
ks.test(plot_frm$logcounts_exp_cv,plot_frm$regulon_activaty_cv)

EC_intersect <- setdiff(score_cst_sig_EC$regulon, exp_cst_sig_EC$regulon)
exp_cst %>% filter(regulon %in% EC_intersect)
#---Smad1
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Smad1",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Smad1", by_exprs_values ="logcounts")

#---Gata6
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Gata6",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Gata6", by_exprs_values ="logcounts")

T1_intersect <- setdiff(score_cst_sig_T1$regulon, c(exp_cst_sig_T1$regulon,exp_cst_sig_T2$regulon))
exp_cst %>% filter(regulon %in% T1_intersect & stage %in% "T1_pre_HSC")
#---Hoxa7
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hif3a",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hif3a", by_exprs_values ="logcounts")

#---Bach1
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Smad3",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Smad3", by_exprs_values ="logcounts")

Adult_intersect <- setdiff(score_cst_sig_Adult$regulon,exp_cst_sig$regulon)
#---Bach1
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Jund",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Jund", by_exprs_values ="logcounts")


FL_intersect <- setdiff(score_cst_sig_FL$regulon,exp_cst_sig$regulon)
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Mecom",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Mecom", by_exprs_values ="logcounts")


#---activate score diff but expression not
#---Ddit3-Adult
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Ddit3",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Ddit3", by_exprs_values ="logcounts")
#---Sox5
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Sox5",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Sox5", by_exprs_values ="logcounts")
#---Batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Batf",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Batf", by_exprs_values ="logcounts")
#---Hoxb5
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hoxb5",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hoxb5", by_exprs_values ="logcounts")
#---Klf5
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Klf5",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Klf5", by_exprs_values ="logcounts")

#---both activate score and expression diff 
#---Gfi1b
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Foxc1",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Foxc1", by_exprs_values ="logcounts")

#---Gfi1b
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Foxc2",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Foxc2", by_exprs_values ="logcounts")

#---Gfi1b
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hoxa4",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hoxa4", by_exprs_values ="logcounts")

#---Gfi1b
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Ebf1",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Ebf1", by_exprs_values ="logcounts")

all_candidates <- c("Smad1","Gata6","Jund","Ddit3","Sox5","Batf","Hoxb5","Klf5","Foxc1","Foxc2","Hoxa4","Ebf1")


```


```{r}
#---Gfi1b
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Gfi1b",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Gfi1b", by_exprs_values ="logcounts")


#---Hoxa2
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hoxa2",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Hoxa2", by_exprs_values ="logcounts")

#---Ikzf1
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Mycn",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Mycn", by_exprs_values ="logcounts")




```
#### 2.8 cell type specific regulon vs expression correlation

```{r}
exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.2_TF_act_sce.rds")
assay(act_sce,"logcounts") <- assay(exp_sce,"logcounts")[row.names(assay(act_sce,"active_score")),]
acticate_mat <- assay(act_sce,"active_score")
exp_mat <- assay(exp_sce,"logcounts")[row.names(acticate_mat),]
cor_p <- sapply(row.names(acticate_mat),function(x,acticate_mat, exp_mat){
  cor(acticate_mat[x,],exp_mat[x,])
},acticate_mat=acticate_mat, exp_mat=exp_mat)
names(cor_p) <- row.names(acticate_mat)

others_cor <- sapply(row.names(acticate_mat),function(x,acticate_mat, exp_mat){
  exp_mat_filter <- exp_mat[setdiff(row.names(exp_mat),x),]
  apply(exp_mat_filter,1,function(y){cor(acticate_mat[x,],y)})
  },acticate_mat=acticate_mat, exp_mat=exp_mat)
others_cor_frm <- data.frame(cor_p=as.vector(others_cor),type=rep("control",length(cor_p)))
plot_frm <- data.frame(cor_p,type=rep("case",length(cor_p)))
plot_frm_all <- rbind(others_cor_frm, plot_frm)

library(ggplot2)
p <- ggplot(plot_frm_all,aes(x=cor_p,color=type,group=type))+
  stat_ecdf()
p

library(ggplot2)
p <- ggplot(plot_frm_all,aes(x=cor_p,fill=type,group=type))+
  geom_density()+
  scale_x_continuous(limits = c(-1,1))
p

```



```{r}
score_cst_sig_EC_kegg <- score_cst_regulons_kegg_sig %>% filter(regulon %in% score_cst_sig_EC$regulon)
stringr::str_to_lower(score_cst_sig_EC_kegg$pathway)
score_cst_sig_EC_kegg$pathway


exp_cst_regulons_kegg <- calcuModuleSimilar(tfr_list = list(EC=exp_cst_sig_EC$regulon), pathway_list = mouse_kegg,adjust = "BH" , nperm=1000, ncores=6)
exp_cst_regulons_kegg_sig <- exp_cst_regulons_kegg %>% filter(padj<0.01)


exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.2_TF_act_sce.rds")

exp_sce_logcounts <- assay(exp_sce,"logcounts")
exp_sce_logcounts_scale <- apply(exp_sce_logcounts,1,function(x){(x-min(x))/(max(x)-min(x))})
exp_sce_logcounts_cv <- apply(exp_sce_logcounts_scale,1,function(x){sd(x)/mean(x)})
act_sce_active_score <- assay(act_sce,"active_score")
act_sce_active_score_scale <- apply(act_sce_active_score,1,function(x){(x-min(x))/(max(x)-min(x))})

act_sce_active_score_cv <- apply(act_sce_active_score_scale,1,function(x){sd(x)/mean(x)})




Reactome_pathway <- fgsea::gmtPathways(gmt.file = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/pathway_data/c2.cp.reactome.v7.4.symbols.gmt")
mouse_reactome <- mapHomologList(input_gene = Reactome_pathway,input_tax = "human",output_tax = "mouse")
sig_regulon_pathway <- calcuModuleSimilar(tfr_list = cst_sig_regulons, pathway_list = mouse_reactome,adjust = "BH" , nperm=1000, ncores=6)
sig_regulon_pathway_sig <- sig_regulon_pathway %>% filter(padj<0.01)

kegg_pathway <- fgsea::gmtPathways(gmt.file = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/pathway_data/c2.cp.kegg.v7.4.symbols.gmt")
mouse_kegg <- mapHomologList(input_gene = kegg_pathway,input_tax = "human",output_tax = "mouse")

kegg_sig_regulon_pathway <- calcuModuleSimilar(tfr_list = cst_sig_regulons, pathway_list = mouse_kegg,adjust = "BH" , nperm=1000, ncores=6)
kegg_sig_regulon_pathway_sig <- kegg_sig_regulon_pathway %>% filter(padj<0.01)

go_term <- fgsea::gmtPathways(gmt.file = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/pathway_data/c5.go.v7.4.symbols.gmt")
mouse_go <- mapHomologList(input_gene = go_term,input_tax = "human",output_tax = "mouse")


go_sig_regulon <- calcuModuleSimilar(tfr_list = cst_sig_regulons, pathway_list = mouse_go,adjust = "BH" , nperm=1000, ncores=6)
go_sig_regulon_sig <- go_sig_regulon %>% filter(padj<0.01)


```




#### 2.7 candidates gene dot plot

```{r}
exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.1_TF_exp_sce.rds")
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/reducedDim/01.2_TF_act_sce.rds")
assay(act_sce,"logcounts") <- assay(exp_sce,"logcounts")[row.names(assay(act_sce,"active_score")),]

exp_cst <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/diff_res/5_stage_expression_level_cst_all_res.txt",sep="\t",header=T)
score_cst <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/diff_res/5_stage_regulon_level_cst_all_res.txt",sep="\t",header=T)

exp_cst_diff <- exp_cst[exp_cst$padj<0.2,]
exp_cst_not_diff <- exp_cst[exp_cst$padj>0.5,]

score_cst_diff <- score_cst[score_cst$padj<0.01,]
score_cst_diff2 <- score_cst_diff[score_cst_diff$CSS>0.4,]
View(score_cst_diff2 %>% filter(!(regulon %in% exp_cst_diff$regulon)))

candidates <- c("Sox17","Myb","Spi1")

tSNE_loc <- reducedDim(act_sce,type = "TSNE")
colnames(tSNE_loc) <- c("x","y")
tSNE_loc$stage <- as.character(colData(act_sce)[row.names(tSNE_loc),"stage"])
exp_tsne_mat <- t(assay(exp_sce,"logcounts")[candidates,])
tSNE_loc <- data.frame(tSNE_loc, exp_tsne_mat[row.names(tSNE_loc),])
tSNE_loc_melt <- reshape2::melt(tSNE_loc,id.vars=c("x","y"),variable.name="gene_name",value.name = "logcounts")

library(ggplot2)
p <- ggplot(data = tSNE_loc_melt,aes(x=x,y=y,))

#---Sox 17
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Sox17",by_exprs_values ="active_score",point_size=1)
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Sox17", by_exprs_values ="logcounts")
#---batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Spi1",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Spi1", by_exprs_values ="logcounts")
#---batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Myb",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Myb", by_exprs_values ="logcounts")

#---batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Tbx3",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Tbx3", by_exprs_values ="logcounts")

#---batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Sox13",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Sox13", by_exprs_values ="logcounts")

#---batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Ppard",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Ppard", by_exprs_values ="logcounts")

#---batf
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Tbx3",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Tbx3", by_exprs_values ="logcounts")
#---Gfi1
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Gfi1",by_exprs_values ="active_score")
plotReducedDim(act_sce, dimred="TSNE",colour_by = "Gfi1", by_exprs_values ="logcounts")

```



#### 2.4 cell type specific TF vs regulon top 500 heatmap

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_test <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/CST_merge_fdr0.01_CSS_0.35.rds")
exp_cst_sig <- cst_test$exp_cst
cst_mat_sig <- cst_test$score_cst
cst_mat_sig$stage <- factor(cst_mat_sig$stage,levels=c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC"))
cst_mat_sig <- cst_mat_sig[order(cst_mat_sig$stage,cst_mat_sig$CSS,decreasing = c(FALSE,TRUE),method = "radix"),]

act_cst_score <- assay(altExp(atfr,"viper"))
sig_act_cst_score <- act_cst_score[unique(cst_mat_sig$regulon),]
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)

anno_col <- as.data.frame(colData(atfr))[,"stage",drop=FALSE]
anno_col$stage <- ifelse(anno_col$stage %in% c("E12","E14"),"FL_HSC",anno_col$stage)
anno_col$stage <- factor(anno_col$stage,levels=c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC"))
anno_row <- tapply(cst_mat_sig$stage,cst_mat_sig$regulon,function(x){paste(x,collapse = "_")})
anno_row <- as.data.frame(anno_row)
colnames(anno_row) <- "stage_type"
anno_row$stage_type <- ifelse(anno_row$stage_type %in% "EC_Adult_HSC","Adult_HSC",anno_row$stage_type)
anno_row$stage_type<- ifelse(anno_row$stage_type %in% "T1_pre_HSC_FL_HSC","T1_pre_HSC",anno_row$stage_type)
levels_stage_type <- c("EC","T1_pre_HSC","FL_HSC","FL_HSC_Adult_HSC","Adult_HSC")
anno_row$stage_type <- factor(anno_row$stage_type,levels=levels_stage_type)
anno_row <- anno_row[order(anno_row$stage_type),,drop=FALSE]
#write.table(anno_row,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_regulon_stage_type.txt",sep="\t",quote=F)
anno_row <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_regulon_stage_type.txt",sep="\t",header=T,row.names=1)
anno_row <- anno_row[,-1,drop=FALSE]

p <- pheatmap(sig_act_cst_score[row.names(anno_row),],scale="row",cluster_cols = F,
              show_rownames = F,treeheight_row = F, clustering_method = "ward.D2",
              cutree_rows = cluster_num, cluster_rows = F, color = pal, 
              show_colnames = F, legend_breaks = seq(-6,6,by=3),
              gaps_col = cumsum(table(anno_col$stage)), annotation_col = anno_col,
              gaps_row = cumsum(table(anno_row$cluster)), annotation_row = anno_row)
p

sig_act_TF_exp <- log2(TF_exp_mat[row.names(anno_row),]+1)
p <- pheatmap(sig_act_TF_exp,scale="row",cluster_cols = F,show_rownames = F,treeheight_row = F,
              clustering_method = "ward.D2",cutree_rows = cluster_num, cluster_rows = F,
              color = pal,show_colnames = F, legend_breaks = seq(-10,10,by=5),
              gaps_col = cumsum(table(anno_col$stage)), annotation_col = anno_col,
              gaps_row = cumsum(table(anno_row$cluster)), annotation_row = anno_row)
p


#---expression level
exp_cst_sig <- cst_test$exp_cst
anno_row <- tapply(exp_cst_sig$stage,exp_cst_sig$regulon,function(x){paste(x,collapse = "_")})
anno_row <- as.data.frame(anno_row)
colnames(anno_row) <- "stage_type"
anno_row$stage_type <- ifelse(anno_row$stage_type %in% "EC_T1_pre_HSC","T1_pre_HSC",anno_row$stage_type)
anno_row$stage_type<- ifelse(anno_row$stage_type %in% c("T2_pre_HSC_FL_HSC"),"FL_HSC",anno_row$stage_type)
levels_stage_type <- c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
anno_row$stage_type <- factor(anno_row$stage_type,levels=levels_stage_type)
anno_row <- anno_row[order(anno_row$stage_type),,drop=FALSE]
write.table(anno_row,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_exp_stage_type.txt",sep="\t",quote=F)
sig_exp_TF_exp <- log2(TF_exp_mat[row.names(anno_row),]+1)
p <- pheatmap(sig_exp_TF_exp,scale="row",cluster_cols = F,show_rownames = F,treeheight_row = F,
              clustering_method = "ward.D2",cutree_rows = cluster_num, cluster_rows = F,
              color = pal,show_colnames = F,legend_breaks = seq(-8,8,by=4),
              gaps_col = cumsum(table(anno_col$stage)), annotation_col = anno_col,
              gaps_row = cumsum(table(anno_row$stage_type)), annotation_row = anno_row)
p


```



#### 2.4 cell type specific TF vs regulon top 500 overlap genes

```{r}
cst_exp_type <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_exp_stage_type.txt",sep="\t",header=T)
act_set_type <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_regulon_stage_type.txt",sep="\t",header=T)

merge_set_frm <- merge(cst_exp_type,act_set_type,by='gene_name')
colnames(merge_set_frm)[2:3] <- c("exp","act")
stage_levels <- c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","FL_HSC_Adult_HSC","Adult_HSC")
merge_set_frm$exp <- factor(merge_set_frm$exp,levels=rev(c("EC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")))
merge_set_frm$act<- factor(merge_set_frm$act,levels=rev(c("EC","T1_pre_HSC","FL_HSC","FL_HSC_Adult_HSC","Adult_HSC")))

library("easyalluvial")
library("parcats")
data <- merge_set_frm[,c(3,5)]
p = alluvial_wide(data, max_variables = 2,fill_by = "last_variable", order_levels=rev(paste("cluster",1:7,sep="")))
parcats(p, marginal_histograms = TRUE, data_input = data)


```




```{r}
TF_act_score_expression_overlap <- intersect(cst_mat_sig$regulon,exp_cst_sig$regulon)
length(setdiff(cst_mat_sig$regulon,exp_cst_sig$regulon))
length(setdiff(exp_cst_sig$regulon,cst_mat_sig$regulon))

#---signicicant activate TF regulons
cst_score <- table(cst_mat_sig$regulon,cst_mat_sig$stage)
attributes(cst_score)$class <- "matrix"
data_set <- as.data.frame(cst_score)
data_set <- data_set[,unique(col_data$stage)]
library(UpSetR)
upset(data_set, nsets = 5, nintersects = 40, mb.ratio = c(0.5, 0.5),
      order.by = c("degree","freq"), decreasing = c(TRUE,TRUE),point.size=5,
      line.size=2.5)
data_set_type <- apply(data_set,1,function(x){paste(names(x[which(x==1)]),collapse = "_")})
table(data_set_type)
levels <- c("EC","EC_T1_pre_HSC","T1_pre_HSC","T1_pre_HSC_T2_pre_HSC","T1_pre_HSC_T2_pre_HSC_Adult_HSC",
            "EC_T1_pre_HSC_T2_pre_HSC","EC_T2_pre_HSC","T2_pre_HSC","T1_pre_HSC_FL_HSC","T1_pre_HSC_T2_pre_HSC_FL_HSC", 
            "T2_pre_HSC_FL_HSC","FL_HSC","EC_Adult_HSC","T2_pre_HSC_Adult_HSC","FL_HSC_Adult_HSC","Adult_HSC")
data_set_type <- data.frame(data_set_type)
data_set_type$data_set_type <- factor(data_set_type$data_set_type,levels=levels)
data_set_type <- data_set_type[order(data_set_type$data_set_type),,drop=FALSE]
#---cst expression type
cst_exp_class <- table(exp_cst_sig$regulon,exp_cst_sig$stage)
attributes(cst_exp_class)$class <- "matrix"
cst_exp_class <- as.data.frame(cst_exp_class)
cst_exp_class <- cst_exp_class[,unique(col_data$stage)]
cst_exp_type <- apply(cst_exp_class,1,function(x){paste(names(x[which(x==1)]),collapse = "_")})
table(cst_exp_type)
levels <- c("EC", "EC_T1_pre_HSC","T1_pre_HSC","T2_pre_HSC","T2_pre_HSC_FL_HSC","FL_HSC","FL_HSC_Adult_HSC","Adult_HSC")
cst_exp_type <- data.frame(cst_exp_type)
cst_exp_type$cst_exp_type <- factor(cst_exp_type$cst_exp_type,levels=levels)
cst_exp_type <- cst_exp_type[order(cst_exp_type$cst_exp_type),,drop=FALSE]
#write.table(cst_exp_type,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/ctsc_exp_stage_type.txt",sep="\t",quote=F)
```

#### 2.6 cell type specific TF vs regulon heatmap

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_exp_type <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_exp_stage_type.txt",sep="\t",header=T)
act_set_type <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_stage_type.txt",sep="\t",header=T)
levels <- intersect(cst_exp_type$stage,act_set_type$stage)
EC_overlap <- lapply(levels,function(x,act,exp){
  intersect(act[act$stage==x,"gene_name"], exp[exp$stage==x,"gene_name"])
}, act=act_set_type, exp=cst_exp_type)
names(EC_overlap) <- levels

library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
col_data <- as.data.frame(colData(atfr))
col_data$stage <- as.character(col_data$stage)
col_data$stage <- ifelse(col_data$stage %in% c("E12","E14"),"FL_HSC",col_data$stage)
col_data_plot <- as.data.frame(col_data[,"stage",drop=FALSE])
col_data_plot$stage <- factor(col_data_plot$stage,levels = unique(col_data_plot$stage))
act_cst_score <- assay(altExp(atfr,"viper"))[row.names(data_set_type ),]
exp_mat <- as.matrix(assay(atfr))

cluster_num <- 6
gene_cluster <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/ctsc_stage_type.txt",sep="\t",header=T)
row.names(gene_cluster) <- gene_cluster$gene_nme
anno_row <- gene_cluster[,3,drop=FALSE]
p <- pheatmap(act_cst_score,scale="row",cluster_cols = F,show_rownames = F,treeheight_row = F,
              clustering_method = "ward.D2",cutree_rows = cluster_num,cluster_rows = F,
              color = pal,show_colnames = F, legend_breaks = seq(-6,6,by=3),
              gaps_row = cumsum(table(gene_cluster$cluster)),annotation_row = anno_row,
              gaps_col = cumsum(table(col_data_plot$stage)), annotation_col = col_data_plot)
p

#---expression level
exp_act_mat <- exp_mat[row.names(gene_cluster),]
exp_act_mat <- log2(exp_act_mat+1)
p <- pheatmap(exp_act_mat,scale="row",cluster_cols = F,show_rownames = F,treeheight_row = F,
              clustering_method = "ward.D2",cutree_rows = cluster_num,cluster_rows = F,
              color = pal,show_colnames = F, legend_breaks = seq(-10,10,by=5),
              gaps_row = cumsum(table(gene_cluster$cluster)),annotation_row = anno_row,
              gaps_col = cumsum(table(col_data_plot$stage)), annotation_col = col_data_plot)
p
cst_mat_sig$cluster <- anno_row[cst_mat_sig$regulon,1]
write.table(cst_mat_sig,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/ctsc_diff_res.txt",sep="\t",row.names=F,quote=F)

#------expression level only heatmap
TF_exp_mat <- exp_mat[cst_exp_type$gene_name,]
exp_act_mat <- log2(TF_exp_mat+1)
anno_row <- cst_exp_type[,3,drop=FALSE]
row.names(anno_row) <- cst_exp_type$gene_name
p <- pheatmap(exp_act_mat,scale="row",cluster_cols = F,show_rownames = F,treeheight_row = F,
              clustering_method = "ward.D2",cluster_rows = F,
              color = pal, show_colnames = F, gaps_row = cumsum(table(anno_row$cluster)),
              annotation_row = anno_row,gaps_col = cumsum(table(col_data_plot$stage)), 
              annotation_col = col_data_plot)
p
```

#### 2.7 cell type specific TF vs regulon candidates plot

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
col_data <- as.data.frame(colData(atfr))
col_data$stage <- as.character(col_data$stage)
col_data$stage <- ifelse(col_data$stage %in% c("E12","E14"),"FL_HSC",col_data$stage)
cst_mat <- findAllCTSRegulons(x = assay(altExp(atfr,"viper")), col_data = col_data,
                              clust_name ="stage",ncores=6, nperm=10000,fdr = 1, css = -1)
cst_mat_sig <- cst_mat[cst_mat$padj<0.01,]
cst_mat_sig2 <- cst_mat_sig[cst_mat_sig$CSS>=0.3,]

TF_exp_mat <- as.matrix(assay(atfr,"counts"))[row.names(assay(altExp(atfr,"viper"))),]
exp_cst <- findAllCTSRegulons(x = TF_exp_mat, col_data = col_data,scale=T,
                              clust_name ="stage",ncores=6, nperm=10000,fdr = 1, css = -1)
exp_cst_sig <- exp_cst[exp_cst$padj<0.01,]
exp_cst_sig2 <- exp_cst_sig[exp_cst_sig$CSS>=0.3,]
#---PCA result
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.2_TF_act_score_PCA_plot.rds")
act_tSNE <- reducedDim(act_sce,"TSNE")
exp_pca <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01_TF_exp_act_score_PCA_plot.rds")
exp_tSNE <- reducedDims(exp_pca[[1]])



```





#### 2.8 cell type specific TF vs regulon sankey plot

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/objects/atfr.rds")
cst_exp_type <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_exp_stage_type.txt",sep="\t",header=T)
act_set_type <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/HSC_development/viper/01.CST_analysis/ctsc_stage_type.txt",sep="\t",header=T)

merge_set_frm <- merge(x = act_set_type,y = cst_exp_type,by="gene_name",all=T)
merge_set_cluster <- merge_set_frm[,c(1,3,5)]
merge_set_cluster <- na.omit(merge_set_cluster)
colnames(merge_set_cluster)[2:3] <- c("Active_score","TF_expression")
plot_data_set <- reshape2::melt(table(merge_set_cluster[,-1]))

library("easyalluvial")
library("parcats")
data <- merge_set_cluster[,c(3:2)]
p = alluvial_wide(data, max_variables = 2,fill_by = "last_variable", order_levels=rev(sort(unique(data$Active_score))))
parcats(p, marginal_histograms = TRUE, data_input = data)


```


#### 1.5 cell type specific genes

```{r}
mat <- as.matrix(logcounts(atfr)[top.gene,])
exp_specific <- findAllCTSRegulons(x = mat, cell_anno=colData(atfr),clust_name="stage")


```

#### 1.6 regulon functions

```{r}
diff_regulons <- as.list(Regulon(atfr)[unique(cst_mat_sig$regulon)])


similarity_regulon <- calcuModuleSimilar(tfr_list = diff_regulons,pathway_list = diff_regulons)


```




