---
title: "02. Cell-type specific regulon analysis during HSC formation"
output: html_notebook
---

#### 1 Introduction

This vignette showed how to calculating the cell-type specific regulons using single-cell RNA sequencing data (Smart-seq2) during HSC formation from Zhou et.al 2018.

##### 1.1 install packages

The scATFR rely on **PIDC** package. So we firstly install these packages.
```{r}
install.packages("devtools")
library(devtools)
devtools::install_github(repo = "hmutpw/PIDC")
devtools::install_github(repo = "hmutpw/scATFR")
```

##### 1.2 loading expression data and create atfr object.

The scATFR accept expression data such as counts, UMIs and TPMs.
```{r}
library(scATFR)
library(PIDC)
exp_data <- readRDS(system.file("./inst/extdata/mouse_HSC_formation_expression.rds",package = "scATFR"))
col_metadata <- readRDS(system.file("./inst/extdata/mouse_HSC_formation_colData.rds",package = "scATFR"))
row.names(col_metadata) <- col_metadata$sample
atfr <- scATFR(exp_data = exp_data, col_data = col_metadata)
```

##### 1.3 Inferring gene regulatory network

```{r}
atfr <- inferGRNs(x = atfr, method="pidc",ncores=6)
```
you can get the regulatory network using the following method:
```{r}
pidc_mat <- GRN(x = atfr, i = "pidc")
```
Also, if you had run you gene regulatory network inferring steps alone, and want to add the network into `atfr` object, you can use `GRN()` methods:
```{r}
pidc_mat <- PIDC(expMat = exp_data, ncores = 6)
GRN(x = atfr,i = "pidc") <- pidc_mat
```

##### 1.4 Inferring regulons from gene regulatory network

```{r}
pidc_net <- matToNet(weightMat = pidc_mat,ncores = 6)
```

#### 2. Integrating the gene regulatory network with local TF-target regulatory information

```{r}
blood_network <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/final_network/mouse/blood_merged_filter_strict_network.txt.gz",sep="\t",header=T)

```









