---
title: "R Notebook"
output: html_notebook
---

##### 1. load data and perform pseudotime analysis

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
expression_matrix <- assay(atfr,'logcounts')
cell_metadata <- colData(atfr)
colData(atfr)$stage <- reLevel(vect = colData(atfr)$stage,level_order = unique(colData(atfr)$stage))
colLabels(atfr) <- colData(atfr)$stage
library(scater)
atfr <- scater::runPCA(x=atfr, ntop =500, exprs_values = "logcounts") 
plotReducedDim(atfr, dimred="PCA",colour_by = "stage",text_by = "stage")
atfr <- scater::runTSNE(x = atfr, ntop =500, exprs_values = "logcounts")
plotReducedDim(atfr, dimred="TSNE",colour_by = "stage",text_by = "stage")
atfr <- scater::runUMAP(x = atfr, exprs_values = "logcounts", ntop =500, pca=50,n_neighbors = 20)
plotReducedDim(atfr, dimred="UMAP",colour_by = "stage",text_by = "stage")

```

##### 1. load data and perform pseudotime analysis with monocle

```{r}
library(monocle)
expression_matrix <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/EHT_gene_count.rds")
col_metadata <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/exp_data/sample2stage.txt",header=T,sep="\t")
row.names(col_metadata) <- col_metadata$sample
col_metadata$stage <- factor(col_metadata$stage,levels=unique(col_metadata$stage))
pd <- new("AnnotatedDataFrame", data = col_metadata)
HSMM <- newCellDataSet(as.matrix(expression_matrix), phenoData = pd)
HSMM <- estimateSizeFactors(HSMM)
HSMM <- estimateDispersions(HSMM)
HSMM <- reduceDimension(HSMM, max_components = 2, num_dim = 6,
                reduction_method = 'tSNE', verbose = T)
HSMM <- clusterCells(HSMM, num_clusters = 2)
plot_cell_clusters(HSMM, 1, 2, color = "stage")

HSMM <- reduceDimension(HSMM, max_components = 2,
    method = 'DDRTree')
HSMM <- orderCells(HSMM)
HSMM$stage <- factor(HSMM$stage,levels=unique(HSMM$stage))

p <- plot_cell_trajectory(HSMM, color_by = "stage",cell_size = 2)+
  scale_color_manual(values=c("#729ece","#ff9e4a","#67bf5c","#ed665d","#ad8bc9","#a8786e"))
ggsave(filename ="pseudotime.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 85, height =100, dpi = 600, limitsize = FALSE, units ="mm")
saveRDS(HSMM,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/HSMM.rds")
col_data <- HSMM@phenoData@data

```

#### 2.1 differential TFs heatmap acorrding pseudotime

```{r}
library(tradeSeq)
library(RColorBrewer)
library(SingleCellExperiment)
library(monocle)
library(dplyr)
cds <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/HSMM.rds")
col_data <- cds@phenoData@data
pseudotime <- col_data[order(col_data$Pseudotime),'Pseudotime',drop=F] %>% as.matrix()
colnames(pseudotime) <- "HSC"
cell_weight <- matrix(1,nrow=nrow(pseudotime))
dimnames(cell_weight) <- list(row.names(pseudotime), "HSC")

x <- pseudotime[,1]
r <- range(x)
b <- seq(from = r[1], to = r[2], length.out = (100))
X <- table(cut(x, breaks = b , include.lowest = TRUE))


stages <- c("AEC","HEC","T1_pre_HSC","T2_pre_HSC","FL_HSC","Adult_HSC")
stage_pseudotime <- lapply(stages,function(i,col_data){
  x <- col_data[,'Pseudotime']
  r <- range(x)
  b <- seq(from = r[1], to = r[2], length.out = 100)
  stage_tbl <- col_data[col_data$stage %in% i,]
  table(cut(stage_tbl$Pseudotime, breaks = b , include.lowest = TRUE))
},col_data=col_data)
names(stage_pseudotime) <- stages
stage_pseudotime_tbl <- do.call(rbind,stage_pseudotime)
#pheatmap(stage_pseudotime_tbl,scale="row",cluster_rows = F,cluster_cols = F)
```


##### 2.1.1 differential genes expression level

```{r}
#---expression level
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
logcounts <- assay(atfr,"logcounts")
trade_sce <- fitGAM(counts = logcounts, cellWeights = cell_weight, pseudotime = pseudotime)
exp_asso_res <- associationTest(trade_sce)
exp_asso_res <- na.omit(exp_asso_res)
exp_asso_res$p.adjust = p.adjust(exp_asso_res$pvalue, method="BH")
saveRDS(exp_asso_res,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/all_gene_diff_test_res.rds")
exp_asso_res_sig <- exp_asso_res[exp_asso_res$p.adjust<0.05,]
exp_asso_res_sig <- exp_asso_res_sig[order(exp_asso_res_sig$p.adjust),]

ysmooth_logcounts <- predictSmooth(models = trade_sce, gene = row.names(logcounts), nPoints = 154)
saveRDS(ysmooth_logcounts,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_all_gene.rds")
saveRDS(ysmooth_logcounts,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_all_gene.rds")

ysmooth_logcounts_list <- split(ysmooth_logcounts,ysmooth_logcounts$gene)
ysmooth_logcounts_list <- lapply(ysmooth_logcounts_list,function(x){
  val <- x[['yhat']]
  names(val) <- x[['time']]
  val
})
ysmooth_logcounts_tbl <- do.call(rbind,ysmooth_logcounts_list)
saveRDS(ysmooth_logcounts_tbl,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_tbl_all_genes.rds")
saveRDS(ysmooth_logcounts_tbl[intersect(row.names(ysmooth_logcounts_tbl),TFs),],"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_tbl-V2.rds")


#---heatmap
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
cluster_num <- 8

p <- pheatmap(ysmooth_logcounts_tbl[row.names(exp_asso_res_sig),],scale="row", cluster_cols = F, clustering_method = "ward.D2", cutree_rows = cluster_num,fontsize = 4, color = pal,show_rownames = F,show_colnames = T,silent=T)
#------how to change cluster order
hclust.result<-p$tree_row
cluster = cutree(hclust.result,k=cluster_num)
cluster_name = paste("C",cluster,sep="")
names(cluster_name) = names(cluster)
hclust.result.order<-p$tree_row$order
cluster_name_ordered = cluster_name[hclust.result.order]
cluster_categroy = unique(cluster_name_ordered)
new_cluster_name = paste("cluster",1:length(cluster_categroy),sep="")
names(new_cluster_name) = cluster_categroy
anno_row = data.frame(cluster = cluster_name_ordered)
anno_row$cluster = new_cluster_name[as.character(anno_row$cluster)]
anno_row$cluster = factor(anno_row$cluster,
levels = paste("cluster",1:cluster_num,sep=""))
anno_col <- col_data[,c("stage","Pseudotime")]
anno_col <- anno_col[order(anno_col$Pseudotime),"stage",drop=F]
row.names(anno_col) <- colnames(plot_mat)
anno_colors = list(stage=c(AEC='#729ece',HEC='#ff9e4a',T1_pre_HSC='#67bf5c',T2_pre_HSC='#ed665d',FL_HSC='#ad8bc9',Adult_HSC='#a8786e'))
TFs <- TF_Symbols$mouse
anno_row_TF <- anno_row[intersect(row.names(anno_row),TFs),,drop=F]

plot_mat <- ysmooth_logcounts_tbl[intersect(row.names(anno_row),TFs),]
p <- pheatmap(plot_mat,scale="row", cluster_cols = F, cluster_rows = F, cutree_rows = cluster_num, fontsize = 4, color = pal, show_rownames = F,show_colnames = T, annotation_row = anno_row_TF, main = "logcounts",annotation_col = anno_col, annotation_colors = anno_colors, gaps_row = cumsum(table(anno_row_TF$cluster)))

ysmooth_logcounts_frm <- data.frame(anno_row_TF, exp_asso_res_sig[row.names(anno_row_TF),],check.names = F)
new_cluster <- c("GP1","GP3","","GP2","GP4","GP6","","GP5")


write.table(ysmooth_logcounts_frm,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm-V2.tsv",sep="\t",quote=F)

ggsave(filename ="differential_all_genes_heatmap_TF_only.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")

median_ysmooth_logcounts <- apply(plot_mat,2,function(x){
  tapply(x,anno_row_TF[names(x),"cluster"],median)
})

median_value_melt <- reshape2::melt(t(median_ysmooth_logcounts))
colnames(median_value_melt) <- c("pseudotime","cluster","logcounts")

library(ggplot2)
q <- ggplot(median_value_melt,aes(x=pseudotime,y=logcounts))+
  geom_smooth()+
 # scale_y_continuous(limits = c(0,8),breaks = seq(0,8,by=4))+
  facet_wrap(.~cluster,scales = "free_y",ncol=1)
q
ggsave(filename ="differential_TF_genes_heatmap_cluster_line-V2.pdf" ,plot = q, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 30, height =200, dpi = 600, limitsize = FALSE, units ="mm")

TF_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm.tsv",sep="\t",header=T)
TF_frm$old_cluster <- ysmooth_logcounts_frm[TF_frm$gene_name,"cluster"]

```

#### 2.2 differential TF regulons heatmap acorrding pseudotime

```{r}
#---regulon activate score level
act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")
active_score_ori <- assay(act_sce,"active_score")
active_score <- t(apply(active_score_ori,1,function(x){10*(x-min(x))/(max(x)-min(x))}))
trade_act <- fitGAM(counts = active_score, cellWeights = cell_weight, pseudotime = pseudotime)
act_asso_res <- associationTest(trade_act)
act_asso_res <- na.omit(act_asso_res)
act_asso_res$p.adjust = p.adjust(act_asso_res$pvalue,method="BH")
act_asso_res_sig <- act_asso_res[act_asso_res$p.adjust<0.05,]
act_asso_res_sig <- act_asso_res_sig[order(act_asso_res_sig$p.adjust),]

plotSmoothers(models = trade_act,active_score,gene="Batf3")
plotSmoothers(models = trade_sce,logcounts,gene="Batf3")

ysmooth_active_score <- predictSmooth(models = trade_act, gene = row.names(active_score), nPoints = 154)
saveRDS(ysmooth_active_score,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score.rds")
ysmooth_active_score_list <- split(ysmooth_active_score,ysmooth_active_score$gene)
ysmooth_active_score_list <- lapply(ysmooth_active_score_list,function(x){
  val <- x[['yhat']]
  names(val) <- x[['time']]
  val
})
ysmooth_active_score_tbl <- do.call(rbind,ysmooth_active_score_list)
saveRDS(ysmooth_active_score_tbl,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score_tbl.rds")
#---heatmap
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
cluster_num <- 8
plot_mat <- ysmooth_active_score_tbl[row.names(act_asso_res_sig),]
p <- pheatmap(plot_mat,scale="row", cluster_cols = F, clustering_method = "ward.D2", cutree_rows = cluster_num,fontsize = 4, color = pal,show_rownames = F,show_colnames = T)
#------how to change cluster order
hclust.result<-p$tree_row
cluster = cutree(hclust.result,k=cluster_num)
cluster_name = paste("C",cluster,sep="")
names(cluster_name) = names(cluster)
hclust.result.order<-p$tree_row$order
cluster_name_ordered = cluster_name[hclust.result.order]
cluster_categroy = unique(cluster_name_ordered)
new_cluster_name = paste("pattern",1:length(cluster_categroy),sep="")
names(new_cluster_name) = cluster_categroy
anno_row = data.frame(cluster = cluster_name_ordered)
anno_row$cluster = new_cluster_name[as.character(anno_row$cluster)]
anno_row$cluster = factor(anno_row$cluster,
levels = paste("pattern",1:cluster_num,sep=""))

p <- pheatmap(plot_mat,scale="row", cluster_cols = F, clustering_method = "ward.D2", cutree_rows = cluster_num,fontsize = 4, color = pal,show_rownames = F,show_colnames = T, annotation_row = anno_row, main = "activate score",annotation_colors = anno_colors)
activate_score_mat <- data.frame(anno_row, act_asso_res_sig[row.names(anno_row),],check.names = F)
write.table(activate_score_frm,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",quote=F)
ggsave(filename ="differential_activate_TF_regulons_heatmap.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")
median_active_score <- apply(plot_mat,2,function(x){
  tapply(x,anno_row[names(x),"pattern"],median)
})


median_value_melt <- reshape2::melt(median_active_score)
colnames(median_value_melt) <- c("cluster","pseudotime","activate_score")
library(ggplot2)
q <- ggplot(median_value_melt,aes(x=pseudotime,y=activate_score))+
  geom_smooth()+
 # scale_y_continuous(limits = c(0,8),breaks = seq(0,8,by=4))+
  facet_wrap(.~cluster,scales = "free_y",ncol=1)
q
ggsave(filename ="differential_activate_TF_regulons_heatmap_cluster_line.pdf" ,plot = q, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 80, height =200, dpi = 600, limitsize = FALSE, units ="mm")

```

#### 2.3 overlap TF and TF regulons according pseudotime

```{r}
logcounts_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm.tsv",sep="\t",header=T,row.names=1)
logcounts_frm <- logcounts_frm[intersect(TFs,row.names(logcounts_frm)),]
activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
overlap_genes <- intersect(row.names(logcounts_frm),row.names(activate_score_frm))

overlap_genes_merged <- data.frame(expression = logcounts_frm[overlap_genes,"pattern"], active_score=activate_score_frm[overlap_genes,"pattern"])
row.names(overlap_genes_merged) <- overlap_genes
library("easyalluvial")
library("parcats")
data <- overlap_genes_merged
p = alluvial_wide(data, max_variables = 2,fill_by = "last_variable", order_levels=rev(paste("pattern",1:8,sep="")))
pa <- parcats(p, marginal_histograms = FALSE, data_input = data)
```

#### 2.4 comparison between TF and TF regulons according pseudotime
```{r}
logcounts_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_tbl.rds")
logcounts_list <- as.list(as_tibble(t(logcounts_mat)))
act_score_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score_tbl.rds")
act_score_list <- as.list(as_tibble(t(act_score_mat)))

logcounts_act_score_cor <- purrr::map2(.x = logcounts_list,.y = act_score_list,.f = function(x,y){
  cor_test <- cor.test(x,y,method = "spearman")
  data.frame(cor=cor_test$estimate,pval=cor_test$p.value)
})
logcounts_act_score_cor_tbl <- do.call(rbind,logcounts_act_score_cor)
logcounts_act_score_cor_tbl$padj <- p.adjust(logcounts_act_score_cor_tbl$pval,method = "BH")
logcounts_act_score_cor_tbl$isSig <- apply(logcounts_act_score_cor_tbl,1,function(x){
  ifelse(x[3]<0.05,"sig","not_sig")
})
saveRDS(logcounts_act_score_cor_tbl,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/logcounts_active_score_spearman.rds")
plot_frm <- logcounts_act_score_cor_tbl
library(ggplot2)
p <- ggplot(data = plot_frm, mapping = aes(x = cor,fill=isSig))+
  geom_histogram(bins = 100)+
  geom_vline(xintercept = c(-0.16,0.16))
p

ggsave(filename ="TF_and_TF_regulons_cor.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 80, height =50, dpi = 600, limitsize = FALSE, units ="mm")
```

#### 2.5 TF and TF regulons patterns correlation

```{r}
logcounts_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm-V2.tsv",sep="\t",header=T,row.names=1)
activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
logcounts_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_tbl-V2.rds")
logcounts_mat <- logcounts_mat[row.names(logcounts_frm),]
act_score_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score_tbl.rds")

logcounts_pattern_mat <- apply(logcounts_mat,2,function(x){tapply(x,logcounts_frm[names(x),"cluster"],median)})
act_score_pattern_mat <- apply(act_score_mat,2,function(x){tapply(x,activate_score_frm[names(x),"cluster"],median)})
act_score_pattern_cor <- cor(t(act_score_pattern_mat),method = "spearman")

cols <- brewer.pal(3, "PRGn")
pal <- colorRampPalette(rev(cols))(100)
p <- pheatmap(act_score_pattern_cor,cluster_rows = T,cluster_cols = T,scale="none",display_numbers = T,color = pal, clustering_method = 'ward.D2', fontsize = 8,legend_breaks = seq(-1,1,by=0.5), breaks = seq(-1,1,length.out=100),number_format = "%.1f",border_color = NA)
ggsave(filename ="act_score_cluster_cor_heatmap.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 90, height =80, dpi = 600, limitsize = FALSE, units ="mm")


cormat <- matrix(NA,nrow=8,ncol=8)
dimnames(cormat) <- list(row.names(logcounts_pattern_mat),row.names(act_score_pattern_mat))
for(i in row.names(logcounts_pattern_mat)){
  for(j in row.names(act_score_pattern_mat)){
    cormat[i,j] <- cor(logcounts_pattern_mat[i,],act_score_pattern_mat[j,],method = "spearman")
  }
}

library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "PRGn")
pal <- colorRampPalette(rev(cols))(100)
p <- pheatmap(cormat,cluster_rows = T,cluster_cols = T,scale="none",display_numbers = T,color = pal, breaks = seq(-1,1,length.out=100), number_format = "%.1f", border_color = NA)
ggsave(filename ="act_score_and_TF_expression_cluster_cor_heatmap-V2.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 100, height =90, dpi = 600, limitsize = FALSE, units ="mm")
```

#### 2.5.2 TF and TF regulons patterns overlap TF genes

```{r}

logcounts_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm-V2.tsv",sep="\t",header=T,row.names=1)
logcounts_all_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_all_diff_gene_frm.tsv",sep="\t",header=T,row.names=1)

activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
overlap_tfs <- intersect(row.names(logcounts_frm),row.names(activate_score_frm))

comparison_frm <- data.frame(logcounts_frm[overlap_tfs,],activate_score_frm[overlap_tfs,])
table(comparison_frm$cluster,comparison_frm$cluster.1)


```

#### 2.6 bulk RNA-seq comparison

```{r}
EHT_bulk <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/EHT_bulk/GSE135600_HSC_GeneLevel_FPKM_5Stages.txt",sep="\t",header=T,row.names = 1)
EHT_bulk <- EHT_bulk[apply(EHT_bulk,1,function(x){length(x[which(x>=1)])>1}),]
gene_infor <- read.table("D:/HSC_Autophagy/ref_genome/gencode.vM25.gene.infor.rm.version.tsv",sep="\t",header=T,row.names=1)

EHT_bulk <- apply(EHT_bulk,2,function(x){tapply(x,as.character(gene_infor[names(x),"gene_name"]),sum)})
stages <- c("EC","EC","HEC","HEC","pre_HSC","pre_HSC","pre_HSC","FL_HSC","FL_HSC","BM_HSC","BM_HSC")
sample2stage <- data.frame(sample=colnames(EHT_bulk),stage=stages)
row.names(sample2stage) <- sample2stage$sample
EHT_bulk_scale <- log2(EHT_bulk+1)


stages_new <- c("AEC","HEC","pre_HSC","FL_HSC","BM_HSC")
EHT_bulk_mean <- t(apply(EHT_bulk,1,function(x){tapply(x,stages,mean)}))
saveRDS(EHT_bulk_mean[,stages_new],"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/EHT_bulk_mean_FPKM.rds")
EHT_bulk_mean <- EHT_bulk_mean[,stages_new]
EHT_bulk_mean <- log2(EHT_bulk_mean+1)

exp_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.1_TF_exp_sce.rds")
TF_exp <- as.matrix(logcounts(exp_sce))
TF_exp_mean <- t(apply(TF_exp,1,function(x){tapply(x,colData(exp_sce)[names(x), "stage"], mean)}))
TF_exp_mean_preHSC <- apply(TF_exp_mean[,3:4],1,mean)
TF_exp_mean <- as.data.frame(TF_exp_mean)
TF_exp_mean$pre_HSC <- TF_exp_mean_preHSC
colnames(TF_exp_mean)[6] <- 'BM_HSC'
TF_exp_mean <- TF_exp_mean[,stages_new]

act_sce <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/viper/reducedDim/01.2_TF_act_sce.rds")
TF_act <- as.matrix(assay(act_sce,"active_score"))
TF_act <- apply(TF_act,1,function(x){(x-min(x))/(max(x)-min(x))})
TF_act_mean <- t(apply(TF_act,2,function(x){tapply(x,colData(act_sce)[names(x), "stage"],mean)}))
TF_act_mean <- as.data.frame(TF_act_mean)
TF_act_mean$pre_HSC <- apply(TF_act_mean[,3:4],1,mean)
colnames(TF_act_mean)[6] <- 'BM_HSC'
TF_act_mean <- TF_act_mean[,stages_new]

overlap_tfs <- intersect(row.names(EHT_bulk_mean),row.names(TF_exp_mean))

bulk_TF_exp <- lapply(overlap_tfs,function(x,bulk,tf_exp){
  cor_res <- cor.test(as.numeric(bulk[x,]),as.numeric(tf_exp[x,]))
  data.frame(gene_name=x, cor = cor_res$estimate, pval = cor_res$p.value)
},bulk=EHT_bulk_mean,tf_exp=TF_exp_mean)
bulk_TF_exp_frm <- do.call(rbind,bulk_TF_exp)
bulk_TF_exp_frm$type <- rep("exp_level",nrow(bulk_TF_exp_frm))


bulk_TF_act <- lapply(overlap_tfs,function(x,bulk,tf_exp){
  cor_res <- cor.test(as.numeric(bulk[x,]),as.numeric(tf_exp[x,]))
  data.frame(gene_name=x, cor = cor_res$estimate, pval = cor_res$p.value)
},bulk=EHT_bulk_mean,tf_exp=TF_act_mean)
bulk_TF_act_frm <- do.call(rbind,bulk_TF_act)
bulk_TF_act_frm$type <- rep("act_level",nrow(bulk_TF_act_frm))

plot_frm <- rbind(bulk_TF_exp_frm, bulk_TF_act_frm)

library(ggplot2)
library(ggridges)
p <- ggplot(plot_frm,mapping = aes(x = cor,fill=type))+
  geom_histogram(bins = 50)
p
p <- ggplot(plot_frm, mapping = aes(x = cor, y= type, fill=type,height=..density..))+
  geom_density_ridges_gradient(size = 1, scale = 1.4, panel_scaling = F, alpha=.8,na.rm=T)

p

library(plyr)
plot_ecdf <- plot_frm
plot_ecdf$cor <- abs(plot_ecdf$cor)
plot_ecdf = ddply(plot_ecdf[,c("type","cor")], .(type), transform, ecdf = ecdf(cor)(cor) )


text_size = 8
line_size = 0.5
p = ggplot(plot_ecdf,aes(x = cor, y=ecdf, color = type))+
stat_ecdf(size = 1)
p

ggsave(filename ="bulk_exp_single_cell_TF_regulons_cor_ecdf.pdf.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 100, height =90, dpi = 600, limitsize = FALSE, units ="mm")

logcounts_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/",sep="\t",header=T,row.names=1)

EHT_kw_test <- pbapply::pbapply(EHT_bulk,1,function(x,stage){
  g_list <- split(x,sample2stage[names(x),]$stage)
  kruskal.test(x = g_list)$p.value
},stage=sample2stage)
EHT_kw_test_padj <- p.adjust(EHT_kw_test,method = "BH")

EHT_anova_test <- pbapply::pbapply(EHT_bulk_scale,1,function(x,sample2stage){
  sample2stage$exp <- x[row.names(sample2stage)]
  anova(lm(formula = exp~stage,data = sample2stage))$`Pr(>F)`[1]
},sample2stage=sample2stage)

EHT_anova_test_padj <- p.adjust(EHT_anova_test,method = "BH")

sample2stage$cell_weight <- rep(1,ncol(EHT_bulk))
sample2stage$pseudotime <- c(1,1.1,2,2.1,3,3.1,3.2,4,4.1,5,5.1)
pseudotime_bulk <- sample2stage[,"pseudotime",drop=FALSE]
cell_weight_bulk <- sample2stage[,"cell_weight",drop=FALSE]
trade_bulk <- fitGAM(counts = EHT_bulk, cellWeights = cell_weight_bulk, pseudotime = pseudotime_bulk)
trade_bulk_res <- associationTest(trade_bulk)
trade_bulk_res <- na.omit(trade_bulk_res)
saveRDS(trade_bulk_res,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/bulk_EHT_diff_test_res.rds")
trade_bulk_res$padj <- p.adjust(trade_bulk_res$pvalue,method = "BH")
trade_bulk_sig <- trade_bulk_res[trade_bulk_res$padj<0.05,]

trade_bulk_sig_tf <- trade_bulk_sig[intersect(row.names(trade_bulk_sig),TFs),]

trade_bulk_TF_regulon_overlap <- intersect(row.names(trade_bulk_sig_tf), row.names(activate_score_frm))
trade_bulk_TFs_overlap <- intersect(row.names(trade_bulk_sig_tf), row.names(logcounts_frm))
trade_bulk_TFs_and_regulons_overlap <- intersect(trade_bulk_TF_regulon_overlap, trade_bulk_TFs_overlap)
TF_and_regulon_overlap <- intersect(row.names(logcounts_frm), row.names(activate_score_frm))

length(setdiff(row.names(trade_bulk_sig_tf),union(row.names(activate_score_frm),row.names(logcounts_frm))))
length(setdiff(row.names(activate_score_frm),union(row.names(trade_bulk_sig_tf),row.names(logcounts_frm))))
length(setdiff(row.names(logcounts_frm),union(row.names(trade_bulk_sig_tf),row.names(activate_score_frm))))

#---heatmap
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
cluster_num <- 8
plot_mat <- EHT_bulk_scale[row.names(trade_bulk_sig_tf),]
p <- pheatmap(plot_mat,scale="row", cluster_cols = F, clustering_method = "ward.D2", cutree_rows = cluster_num,fontsize = 4, color = pal,show_rownames = F,show_colnames = T)
#------how to change cluster order
hclust.result<-p$tree_row
cluster = cutree(hclust.result,k=cluster_num)
cluster_name = paste("C",cluster,sep="")
names(cluster_name) = names(cluster)
hclust.result.order<-p$tree_row$order
cluster_name_ordered = cluster_name[hclust.result.order]
cluster_categroy = unique(cluster_name_ordered)
new_cluster_name = paste("pattern",1:length(cluster_categroy),sep="")
names(new_cluster_name) = cluster_categroy
anno_row = data.frame(cluster = cluster_name_ordered)
anno_row$cluster = new_cluster_name[as.character(anno_row$cluster)]
anno_row$cluster = factor(anno_row$cluster,
levels = paste("pattern",1:cluster_num,sep=""))

p <- pheatmap(plot_mat,scale="row", cluster_cols = F, clustering_method = "ward.D2", cutree_rows = cluster_num,fontsize = 4, color = pal,show_rownames = F,show_colnames = T, annotation_row = anno_row, main = "activate score",annotation_colors = anno_colors)
activate_score_mat <- data.frame(anno_row, EHT_bulk_scale[row.names(anno_row),],check.names = F)
write.table(activate_score_frm,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm_bulk_seq.tsv",sep="\t",quote=F)

ggsave(filename ="differential_TF_gene_bulk_seq__heatmap.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")
median_active_score <- apply(plot_mat,2,function(x){
  tapply(x,anno_row[names(x),"pattern"],median)
})







```

##### 2.6.1 bulk clusters heatmap

```{r}
stages_new <- c("AEC","HEC","pre_HSC","FL_HSC","BM_HSC")
EHT_bulk_mean <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/EHT_bulk_mean_FPKM.rds")

trade_bulk_res <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/bulk_EHT_diff_test_res.rds")
trade_bulk_diff <- trade_bulk_res[trade_bulk_res$padj<0.05,]

trade_bulk_diff_fpkm <- EHT_bulk_mean[row.names(trade_bulk_diff),]

#---heatmap
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
cluster_num <- 8
plot_mat <- log2(trade_bulk_diff_fpkm+1)
p <- pheatmap(plot_mat, scale="row", cluster_cols = F, clustering_method = "ward.D2", cutree_rows = cluster_num,fontsize = 4, color = pal,show_rownames = F,show_colnames = T)
#------how to change cluster order
hclust.result<-p$tree_row
cluster = cutree(hclust.result,k=cluster_num)
cluster_name = paste("C",cluster,sep="")
names(cluster_name) = names(cluster)
hclust.result.order<-p$tree_row$order
cluster_name_ordered = cluster_name[hclust.result.order]
cluster_categroy = unique(cluster_name_ordered)
new_cluster_name = paste("pattern",1:length(cluster_categroy),sep="")
names(new_cluster_name) = cluster_categroy
anno_row = data.frame(cluster = cluster_name_ordered)
anno_row$cluster = new_cluster_name[as.character(anno_row$cluster)]
anno_row$cluster = factor(anno_row$cluster,
levels = paste("pattern",1:cluster_num,sep=""))

trade_bulk_diff_sorted <- data.frame(gene_name = row.names(anno_row),anno_row, trade_bulk_diff[row.names(anno_row),])
write.table(trade_bulk_diff_sorted,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/EHT_diff_frm_bulk_seq.tsv",sep="\t",row.names = F, quote=F)

anno_row_tf <- anno_row[intersect(row.names(anno_row),TF_Symbols$mouse),,drop=FALSE]
trade_bulk_diff_tf_sorted <- trade_bulk_diff_sorted[intersect(row.names(trade_bulk_diff_sorted),TF_Symbols$mouse),]
write.table(trade_bulk_diff_tf_sorted,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/EHT_diff_TFs_from_bulk_seq.tsv",sep="\t",row.names = F, quote=F)


plot_tf_mat <- plot_mat[row.names(anno_row_tf),]

p <- pheatmap(plot_tf_mat, scale="row", cluster_cols = F,cluster_rows = F, fontsize = 4, color = pal, show_rownames = F,show_colnames = T, annotation_row = anno_row, main = "TF",annotation_colors = anno_colors,gaps_row = cumsum(table(anno_row_tf$cluster)))

ggsave(filename ="differential_TF_genes_heatmap_bulk_RNA-seq.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")


mean_tf_exp <- apply(plot_tf_mat,2,function(x){tapply(x,anno_row_tf[names(x),'cluster'], mean)})

mean_tf_exp_melt <- reshape2::melt(mean_tf_exp)
mean_tf_exp_melt <- reshape2::melt(plot_tf_mat)
mean_tf_exp_melt$group <- anno_row_tf[mean_tf_exp_melt$cluster,"cluster"]
colnames(mean_tf_exp_melt) <- c("cluster","stage","logFPKM")
#------get my theme
text_size = 8
line_size = 0.5/1.07
library(ggplot2)
#------set plot theme
mytheme = theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA,color=NA),
strip.background = element_rect(fill=NA,color=NA),
strip.text = element_text(size=text_size,color="black"),
panel.grid.major = element_line(size=line_size/2, linetype = 8, color = "lightgray"),
panel.grid.minor = element_line(size=NA, color = NA),
axis.ticks = element_line(size=line_size,color="black"),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45, hjust = 1, vjust =1,color="black"), 
axis.text.y= element_text(size=text_size,color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

plot_frm <- mean_tf_exp_melt
q <- ggplot(plot_frm, aes(x=stage,y=logFPKM,group=cluster))+
  geom_line(color="red")+
  geom_point(shape=16,color="red")+
#  geom_smooth(method="loess",color="red")+
  facet_wrap(.~cluster, scale="free_y",ncol=1)+
  mytheme
q





```

##### 2.6.2 functional_enrichment

```{r}
trade_bulk_diff_sorted <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/EHT_diff_frm_bulk_seq.tsv",sep="\t",header=T)
row.names(trade_bulk_diff_sorted) <- trade_bulk_diff_sorted$gene_name

trade_bulk_diff_list <- split(trade_bulk_diff_sorted$gene_name,trade_bulk_diff_sorted$cluster)
reactome_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/scATFR/inst/extdata/Reactome_pathway.rds")
mouse_reactome <- reactome_pathway$`Mus musculus`
mouse_reactome_num <- sapply(mouse_reactome,length)
mouse_reactome_filter <- mouse_reactome[between(mouse_reactome_num,10,500)]

trade_bulk_diff_cluster_enrichment <- calcuModuleSimilar(tfr_list = trade_bulk_diff_list, pathway_list = mouse_reactome_filter, adjust = "BH" , ncores=6, fdr = 0.05, min_overlap = 5)
fwrite(trade_bulk_diff_cluster_enrichment,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/EHT_bulk_seq_each_pattern_reactome_enrichment.tsv",sep="\t")

```


#### 2.7 TF and TF regulons candidates expression during pseudotime

```{r}
logcounts_frm <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts.rds")
act_score_frm <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score.rds")

ysmooth_logcounts_act_score <- full_join(logcounts_frm, act_score_frm,by=c('lineage'='lineage','time'='time', 'gene'='gene'))
colnames(ysmooth_logcounts_act_score)[4:5] <- c("logcounts", "active_score")

logcounts_active_score_spearman <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/logcounts_active_score_spearman.rds")
logcounts_active_score_negative <- logcounts_active_score_spearman %>% filter(cor < -0.5 & padj<0.05)

diff_act_res <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T)

diff_act_neg_gene <- intersect(row.names(diff_act_res),row.names(logcounts_active_score_negative))
diff_act_neg_tab <- diff_act_res[diff_act_neg_gene,]
diff_act_neg_tab$cor <- logcounts_active_score_negative[diff_act_neg_gene,"cor"]


i <- c("Hoxa9","Mycn","Gata2","Sox17","Sox7","Meis2","Myb","Bcl11a","Spi1","Runx1","Hlf","Tbx3","Foxo4","Hmga1","Dnmt1","Hoxa2","Hoxa3","Hoxa4","Hoxb5","Satb2","Sox2","Bptf","Ncoa2","Snai1","Cebpa","Sox2","Nfkb2","Cebpb","Erg","Zzz3","Bbx","Ngfa","Ash1l","Zkscan3","Ubp1","Klf6", "Nfix","Zbtb7a","Prdm16","Sp3")

i <- c("Twist1")

#------get my theme
text_size = 8
line_size = 0.5/1.07
library(ggplot2)
#------set plot theme
mytheme = theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA,color=NA),
strip.background = element_rect(fill=NA,color=NA),
strip.text = element_text(size=text_size,color="black"),
panel.grid.major = element_line(size=line_size/2, linetype = 8, color = "lightgray"),
panel.grid.minor = element_line(size=NA, color = NA),
axis.ticks = element_line(size=line_size,color="black"),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45, hjust = 1, vjust =1,color="black"), 
axis.text.y= element_text(size=text_size,color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')


plot_frm <- ysmooth_logcounts_act_score %>% filter(gene %in% i)


p <- ggplot(plot_frm,aes(x=time))+
  geom_smooth(mapping = aes(y=active_score),method="loess",color="red")+
  facet_wrap(.~gene,scale="free")+
  mytheme
p
ggsave(filename ="act_score_cluster_candidates.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 200, height =200, dpi = 600, limitsize = FALSE, units ="mm")
q <- ggplot(plot_frm,aes(x=time))+
  geom_smooth(mapping = aes(y=logcounts),method="loess",color="blue")+
  facet_wrap(.~gene,scale="free")+
  mytheme
q
p+q
ggsave(filename ="logcounts_cluster_candidates.pdf" ,plot = q, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 210, height =200, dpi = 600, limitsize = FALSE, units ="mm")
#---not used 
scaleFactor <- max(plot_frm$logcounts) / max(plot_frm$active_score)
scaleFactor <- 2
library(ggplot2)
q <- ggplot(plot_frm,aes(x=time))+
  geom_smooth(mapping = aes(y=logcounts),method="loess",color="blue")+
  geom_smooth(mapping = aes(y=active_score*scaleFactor),method="loess",color="red")+
  scale_y_continuous(
    # Features of the first axis
    name = "logcounts",
    sec.axis = sec_axis(~./scaleFactor, name="active_score")
  )+
  facet_wrap(.~gene)
q
```


#### 2.7 TF regulons similarity correlation analysis

```{r}
atfr_regulons <- as.list(Regulon(atfr))
atfr_regulon_similarity <- calcuModuleSimilar(tfr_list = atfr_regulons, pathway_list = atfr_regulons,fdr = 2,min_overlap = 0,ncores = 6, collapse_set = F)
atfr_regulon_similarity_filter <- atfr_regulon_similarity %>% filter(similarity >0 & regulon != pathway)

activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
act_score_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score_tbl.rds")
act_score_mat_cor <- cor(t(act_score_mat),method = "spearman")
activate_score_frm_c1 <- activate_score_frm %>% filter(cluster %in% "cluster1")
diff_score_tf <- atfr_regulon_similarity_filter %>% filter(regulon %in% row.names(activate_score_frm_c1) & pathway %in% row.names(activate_score_frm_c1))
diff_score_tf_int <- diff_score_tf[,1:2]
diff_score_tf_int <- unique(as.data.frame(t(apply(diff_score_tf_int,1,function(x){sort(x)}))))
write.table(diff_score_tf_int,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/diff_activate_score_similarity_network.tsv",sep="\t",row.names = F,quote=F)
diff_score_tf_mat <- reshape2::dcast(atfr_regulon_similarity_filter[,1:3],formula = regulon~pathway)
row.names(diff_score_tf_mat) <- diff_score_tf_mat$regulon
diff_score_tf_mat <- diff_score_tf_mat[,-1]
diff_score_tf_mat[is.na(diff_score_tf_mat)] <- 0

act_score_mat_cor_norm <- sqrt(diff_score_tf_mat*abs(act_score_mat_cor))

diff_score_tf_mat_module <- PIDC::matToNet(act_score_mat_cor_norm,cutoff = 0,ncores = 4)
diff_score_tf_mat_module <- diff_score_tf_mat_module %>% filter(regulator %in% row.names(activate_score_frm) & target %in% row.names(activate_score_frm))
diff_score_tf_int <- unique(as.data.frame(t(apply(diff_score_tf_mat_module[,1:2],1,function(x){sort(x)}))))

write.table(diff_score_tf_int,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/diff_activate_score_similarity_network.tsv",sep="\t",row.names = F,quote=F)

anno_row <- activate_score_frm[,1,drop=F]
pheatmap(diff_score_tf_mat,scale="none",clustering_method = "ward.D2", fontsize = 3,annotation_row = anno_row)

```
#### 2.8 TF regulons TF expression correlation in each cluster

```{r}
activate_score_cluster <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
act_score_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score_tbl.rds")
tf_exp_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_tbl-V2.rds")

overlap_tfs <- intersect(row.names(activate_score_cluster), row.names(act_score_mat))

TF_regulon_cor <- lapply(overlap_tfs,function(x,tf_exp_mat,act_score_mat){
  cor_res <- cor.test(x = as.numeric(tf_exp_mat[x,]),y=as.numeric(act_score_mat[x,]),method="spearman")
  data.frame(gene_name=x,cor=cor_res$estimate,pval=cor_res$p.value)
},act_score_mat=act_score_mat, tf_exp_mat=tf_exp_mat)
TF_regulon_cor_tbl <- do.call(rbind,TF_regulon_cor)
row.names(TF_regulon_cor_tbl) <- TF_regulon_cor_tbl$gene_name
TF_regulon_cor_tbl$cluster <- activate_score_cluster[TF_regulon_cor_tbl$gene_name,"cluster"]

library(ggplot2)
library(ggridges)
text_size = 8
line_size = 0.5
mytheme <- theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA),
panel.grid.major = element_line(size=0.25*line_size, linetype = 8, color = "lightgray"),
axis.ticks = element_line(size=line_size),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, color="black"), 
axis.text.y= element_text(size=text_size, color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

plot_frm <- TF_regulon_cor_tbl
plot_frm$cluster <- factor(plot_frm$cluster,levels=rev(sort(unique(plot_frm$cluster))))

p <- ggplot(data = plot_frm,mapping = aes(x=cor, y=cluster, fill=cluster,height=..density..))+
  geom_density_ridges_gradient(scale = 1.4, panel_scaling = F, alpha=.8,na.rm=T)+
  labs(x='cor',y='',title = '')+
  scale_x_continuous(limits = c(-1,1))+
  mytheme
p

ggsave(filename ="act_score_tf_expression_correlation_by_cluster.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 100, height =120, dpi = 600, limitsize = FALSE, units ="mm")

library(plyr)
plot_ecdf = ddply(plot_frm[,c("cluster","cor")], .(cluster), transform, ecdf = ecdf(cor)(cor) )
q = ggplot(plot_ecdf, aes(x = cor, y=ecdf, color = cluster))+
stat_ecdf(size = line_size)+
mytheme
q


```



#### 3 pathway enrichment analysis

#### 3.1 reactome pathway in different patterns

```{r}
#---pathways similarity analysis
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/objects/atfr.rds")
activate_score_regulon <- as.list(Regulon(atfr)[row.names(activate_score_frm)])
reactome_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/scATFR/inst/extdata/Reactome_pathway.rds")
mouse_reactome <- reactome_pathway$`Mus musculus`
mouse_reactome_num <- sapply(mouse_reactome,length)
mouse_reactome_filter <- mouse_reactome[between(mouse_reactome_num,10,500)]

score_cst_regulons_reactome <- calcuModuleSimilar(tfr_list = activate_score_regulon, pathway_list = mouse_reactome_filter, adjust = "BH" , ncores=6, fdr = 0.05, min_overlap = 5)
score_cst_regulons_reactome$pattern <- activate_score_frm[score_cst_regulons_reactome$regulon,"cluster"]
saveRDS(score_cst_regulons_reactome,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/All_8_patterns_reactome_enrichment_res.rds")
fwrite(score_cst_regulons_reactome,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/All_8_patterns_reactome_enrichment_res.tsv",sep="\t")

score_cst_regulons_reactome_list <- split(score_cst_regulons_reactome$pathway,score_cst_regulons_reactome$regulon)
overlp_term_frm <- calcuModuleSimilar(tfr_list = score_cst_regulons_reactome_list, pathway_list = score_cst_regulons_reactome_list, adjust = "BH" , ncores=6, fdr = 0.05, min_overlap = 1,collapse_set = F)
overlp_term_frm_sig <- overlp_term_frm %>% filter(padj<0.05) %>% filter(regulon != pathway)
overlp_term_frm_sig_int <- as.data.frame(t(apply(overlp_term_frm_sig[,1:2],1,function(x){sort(x)})))
overlp_term_frm_sig_int <- cbind(overlp_term_frm_sig_int,overlp_term_frm_sig[,3:4])
overlp_term_frm_sig_int <- unique(overlp_term_frm_sig_int)

overlp_term_mat <- matrix(0, nrow=length(score_cst_regulons_reactome_list),ncol=length(score_cst_regulons_reactome_list))
dimnames(overlp_term_mat) <- list(names(score_cst_regulons_reactome_list), names(score_cst_regulons_reactome_list))
for(i in names(score_cst_regulons_reactome_list)){
  for(j in names(score_cst_regulons_reactome_list)){
    overlp_term_mat[i,j] <- length(intersect(score_cst_regulons_reactome_list[[i]],score_cst_regulons_reactome_list[[j]]))
    
  }
}
diag(overlp_term_mat) <- 0
overlp_term_mat_melt <- reshape2::melt(overlp_term_mat) %>% filter(value>0)
overlp_term_mat_melt <- as.data.frame(t(apply(overlp_term_mat_melt,1,function(x){c(sort(x[1:2]),x[3])}))) %>% unique()
write.table(overlp_term_frm_sig_int,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/diff_activate_score_similarity_network.tsv",sep="\t",row.names = F,quote=F)
activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
#---1
module1_tfs <- read.csv("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/mcode/module1.csv")
module1_tfs <- unique(module1_tfs$name)
module1_tfs_mat <- overlp_term_mat[module1_tfs,module1_tfs]
module1_tfs_reactome <- score_cst_regulons_reactome %>% filter(regulon %in% module1_tfs)
head(sort(table(module1_tfs_reactome$pathway),decreasing = T))
module1_tfs_cluster <- table(activate_score_frm[module1_tfs,"cluster"])

#---2
module2_tfs <- read.csv("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/mcode/module2.csv")
module2_tfs <- module2_tfs$name
module2_tfs_reactome <- score_cst_regulons_reactome %>% filter(regulon %in% module2_tfs)
head(sort(table(module2_tfs_reactome$pathway),decreasing = T))
module2_tfs_cluster <- table(activate_score_frm[module2_tfs,"cluster"])
pie(module2_tfs_cluster)
#---3
module3_tfs <- read.csv("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/mcode/module3.csv")
module3_tfs <- module3_tfs$name
module3_tfs_reactome <- score_cst_regulons_reactome %>% filter(regulon %in% module3_tfs)
head(sort(table(module3_tfs_reactome$pathway),decreasing = T))
module3_tfs_cluster <- table(activate_score_frm[module3_tfs,"cluster"])
pie(module3_tfs_cluster)

#---4
module4_tfs <- read.csv("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/mcode/module4.csv")
module4_tfs <- module4_tfs$name
module4_tfs_reactome <- score_cst_regulons_reactome %>% filter(regulon %in% module4_tfs)
head(sort(table(module4_tfs_reactome$pathway),decreasing = T))
module4_tfs_cluster <- table(activate_score_frm[module4_tfs,"cluster"])
pie(module4_tfs_cluster)

#---5
module5_tfs <- read.csv("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/mcode/module5.csv")
module5_tfs <- module5_tfs$name
module5_tfs_reactome <- score_cst_regulons_reactome %>% filter(regulon %in% module5_tfs)
head(sort(table(module5_tfs_reactome$pathway),decreasing = T))
module5_tfs_cluster <- table(activate_score_frm[module5_tfs,"cluster"])
pie(module5_tfs_cluster)

#---6
module6_tfs <- read.csv("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/network/mcode/module6.csv")
module6_tfs <- module6_tfs$name
module6_tfs_reactome <- score_cst_regulons_reactome %>% filter(regulon %in% module6_tfs)
head(sort(table(module6_tfs_reactome$pathway),decreasing = T))
module6_tfs_cluster <- table(activate_score_frm[module6_tfs,"cluster"])
pie(module6_tfs_cluster)
```

#### 3.2 reactome pathway in different patterns heatmap

```{r}
#----
activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)
activate_score_frm$gene <- row.names(activate_score_frm)
activate_score_frm_new <- activate_score_frm %>% tidyr::separate_rows(DBD,sep="; ")
#---Module1
module1_activate <- activate_score_frm_new %>%filter(gene %in% module1_tfs)
table(module1_activate$DBD)/table(activate_score_frm_new$DBD)[names(table(module1_activate$DBD))]
module2_activate <- activate_score_frm_new %>%filter(gene %in% module2_tfs)
table(module2_activate$DBD)/table(activate_score_frm_new$DBD)[names(table(module2_activate$DBD))]
table(module2_activate$DBD)
module3_activate <- activate_score_frm_new %>%filter(gene %in% module3_tfs)
table(module3_activate$DBD)/table(activate_score_frm_new$DBD)[names(table(module3_activate$DBD))]
table(module3_activate$DBD)
module4_activate <- activate_score_frm_new %>%filter(gene %in% module4_tfs)
table(module4_activate$DBD)/table(activate_score_frm_new$DBD)[names(table(module4_activate$DBD))]
table(module4_activate$DBD)
module5_activate <- activate_score_frm_new %>%filter(gene %in% module5_tfs)
table(module5_activate$DBD)/table(activate_score_frm_new$DBD)[names(table(module5_activate$DBD))]
table(module5_activate$DBD)
module6_activate <- activate_score_frm_new %>%filter(gene %in% module6_tfs)
table(module6_activate$DBD)/table(activate_score_frm_new$DBD)[names(table(module6_activate$DBD))]
table(module6_activate$DBD)

module1_tfs

score_cst_regulons_reactome_filter <- score_cst_regulons_reactome %>% filter(firstInGroup==1)

score_cst_regulons_reactome_clusters <- table(score_cst_regulons_reactome_filter$pathway,score_cst_regulons_reactome_filter$pattern)

plot_mat <- score_cst_regulons_reactome_clusters
plot_mat <- plot_mat[rowSums(plot_mat)>=2,]

library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
cluster_num <- 8




anno_row <- rowSums(table(score_cst_regulons_reactome_filter$pathway,score_cst_regulons_reactome_filter$regulon))
anno_row <- as.data.frame(anno_row[row.names(plot_mat)])
colnames(anno_row) <- "num"
p <- pheatmap(plot_mat,scale="row",show_rownames = T, fontsize = 6,clustering_method = "ward.D2", cluster_cols = F, color=pal, annotation_row = anno_row, treeheight_row = F)

ggsave(filename ="reactome_pathway_cluster.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/",
       width = 200, height =250, dpi = 600, limitsize = FALSE, units ="mm")



```

###3.3 reactome enrichment each cluster

```{r}
score_cst_regulons_reactome <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/All_8_patterns_reactome_enrichment_res.tsv",sep="\t",data.table=F)

#---cluster1
score_cst_regulons_reactome_c1 <- score_cst_regulons_reactome %>% filter(pattern=="cluster1" & firstInGroup==1)
score_cst_regulons_reactome_c1_top <- head(sort(table(score_cst_regulons_reactome_c1$pathway),decreasing = T),10)
selected_terms1 <- names(score_cst_regulons_reactome_c1_top)[1:4]
score_cst_regulons_reactome_c1_selected <- score_cst_regulons_reactome_c1 %>% filter(pathway %in% selected_terms)

overlp_C1_tf_num <- sapply(score_cst_regulons_reactome_c1_list,length)
overlp_C1_tf_num <- data.frame(gene=names(overlp_C1_tf_num),num = overlp_C1_tf_num)
write.table(overlp_C1_tf_num,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/network/cluster1_node_attr.tsv",sep="\t",quote=F,row.names=F)
#---cluster2
score_cst_regulons_reactome_c2 <- score_cst_regulons_reactome %>% filter(pattern=="cluster2")
score_cst_regulons_reactome_c2_top <- head(sort(table(score_cst_regulons_reactome_c2$pathway),decreasing = T))
selected_terms2 <- c('RHO GTPase cycle','Nuclear Receptor transcription pathway', 'MAPK family signaling cascades')
score_cst_regulons_reactome_c1_selected <- score_cst_regulons_reactome_c1 %>% filter(pathway %in% selected_terms)
#---cluster3
score_cst_regulons_reactome_c3 <- score_cst_regulons_reactome %>% filter(pattern=="cluster3" & firstInGroup==1)
score_cst_regulons_reactome_c3_top <- head(sort(table(score_cst_regulons_reactome_c3$pathway),decreasing = T))

selected_terms3 <- names(score_cst_regulons_reactome_c3_top)[1:4]

#---cluster4
score_cst_regulons_reactome_c4 <- score_cst_regulons_reactome %>% filter(pattern=="cluster4" & firstInGroup==1)
score_cst_regulons_reactome_c4_top <- head(sort(table(score_cst_regulons_reactome_c4$pathway),decreasing = T))
selected_terms4 <- names(score_cst_regulons_reactome_c4_top)[1:4]

#---cluster5
score_cst_regulons_reactome_c5 <- score_cst_regulons_reactome %>% filter(pattern=="cluster5" & firstInGroup==1)
score_cst_regulons_reactome_c5_top <- head(sort(table(score_cst_regulons_reactome_c5$pathway),decreasing = T))
selected_terms5 <- c('Cell Cycle, Mitotic',"Regulation of TP53 Activity",'RHO GTPase cycle',"Signaling by Nuclear Receptors")

#---cluster6
score_cst_regulons_reactome_c6 <- score_cst_regulons_reactome %>% filter(pattern=="cluster6" & firstInGroup==1)
score_cst_regulons_reactome_c6_top <- head(sort(table(score_cst_regulons_reactome_c6$pathway),decreasing = T))
selected_terms6 <- c('Cytokine Signaling in Immune system','Cellular responses to stress','Signaling by PDGF','SUMOylation')

#---cluster7
score_cst_regulons_reactome_c7 <- score_cst_regulons_reactome %>% filter(pattern %in% c("cluster7")& firstInGroup==1)
score_cst_regulons_reactome_c7_top <- head(sort(table(score_cst_regulons_reactome_c7$pathway),decreasing = T))
selected_terms7 <- names(score_cst_regulons_reactome_c7_top)[1:2]

#---cluster8
score_cst_regulons_reactome_c8 <- score_cst_regulons_reactome %>% filter(pattern=="cluster8"& firstInGroup==1)
score_cst_regulons_reactome_c8_top <- head(sort(table(score_cst_regulons_reactome_c8$pathway),decreasing = T))
selected_terms8 <- names(score_cst_regulons_reactome_c8_top)[1:2]

merged_terms <- unique(c(selected_terms1, selected_terms2, selected_terms3, selected_terms4, selected_terms5, selected_terms6, selected_terms7, selected_terms8))

#---
selected_reactome_term <- score_cst_regulons_reactome %>% filter(firstInGroup==1 & pathway %in% merged_terms)
selected_reactome_term_stat <- table(selected_reactome_term$pathway,selected_reactome_term$pattern)




plot_frm <- reshape2::melt(selected_reactome_term_stat)
colnames(plot_frm) <- c("pathway","cluster","number")
plot_frm <- plot_frm[plot_frm$number>0,]
plot_frm$pathway <- factor(plot_frm$pathway,levels=rev(merged_terms))
#------get my theme
text_size = 8
line_size = 0.5/1.07
library(ggplot2)
#------set plot theme
mytheme = theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA,color=NA),
strip.background = element_rect(fill=NA,color=NA),
strip.text = element_text(size=text_size,color="black"),
panel.grid.major = element_line(size=line_size/2, linetype = 8, color = "lightgray"),
panel.grid.minor = element_line(size=NA, color = NA),
axis.ticks = element_line(size=line_size,color="black"),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45, hjust = 1, vjust =1,color="black"), 
axis.text.y= element_text(size=text_size,color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')


p <- ggplot(plot_frm,aes(x=cluster,y=pathway,size=number))+
  geom_point(shape=16)+
  scale_y_discrete(position = "right")+
  mytheme
p
ggsave(filename ="enrichment_terms.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/",
       width = 230, height =100, dpi = 600, limitsize = FALSE, units ="mm")
```
#### 3.4 all diff TF vs regulon level enrichment res

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/")
logcounts_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_logcounts_tbl_all_genes.rds")
logcounts_mat_mat <- 
all_test_gene <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/all_gene_diff_test_res.rds")
all_diff_gene <- all_test_gene %>% filter(p.adjust<0.05)
logcounts_diff_mat <- logcounts_mat[row.names(all_diff_gene),]



library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
cluster_num <- 8
p <- pheatmap(mat = logcounts_diff_mat,scale = "row",show_rownames = F,show_colnames = F,color = pal,cutree_rows = cluster_num, silent = T,clustering_method = "ward.D2")

#------how to change cluster order
hclust.result<-p$tree_row
cluster = cutree(hclust.result,k=cluster_num)
cluster_name = paste("C",cluster,sep="")
names(cluster_name) = names(cluster)
hclust.result.order<-p$tree_row$order
cluster_name_ordered = cluster_name[hclust.result.order]
cluster_categroy = unique(cluster_name_ordered)
new_cluster_name = paste("pattern",1:length(cluster_categroy),sep="")
names(new_cluster_name) = cluster_categroy
anno_row = data.frame(cluster = cluster_name_ordered)
anno_row$cluster = new_cluster_name[as.character(anno_row$cluster)]
anno_row$cluster = factor(anno_row$cluster,
levels = paste("pattern",1:cluster_num,sep=""))

#---anno col
cds <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/HSMM.rds")
col_data <- cds@phenoData@data
anno_col <- col_data [,c("stage","Pseudotime")]
anno_col <- anno_col[order(anno_col$Pseudotime),"stage",drop=F]
row.names(anno_col) <- colnames(plot_mat)
TFs <- TF_Symbols$mouse
anno_row_TF <- anno_row[intersect(row.names(anno_row),TFs),,drop=F]
anno_colors = list(stage=c(AEC='#729ece',HEC='#ff9e4a',T1_pre_HSC='#67bf5c',T2_pre_HSC='#ed665d',FL_HSC='#ad8bc9',Adult_HSC='#a8786e'))


plot_mat <- logcounts_diff_mat[row.names(anno_row_TF),]
p <- pheatmap(plot_mat,scale="row", cluster_cols = F, cluster_rows = F, fontsize = 4, color = pal,show_rownames = F,show_colnames = T, annotation_row = anno_row_TF, annotation_col = anno_col, main = "logcounts",annotation_colors = anno_colors, gaps_row = cumsum(table(anno_row_TF$cluster)))


ysmooth_logcounts_frm <- data.frame(anno_row, exp_asso_res_sig[row.names(anno_row),],check.names = F)
write.table(ysmooth_logcounts_frm,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm-V2.tsv",sep="\t",quote=F)

ggsave(filename ="differential_TF_genes_heatmap-V2.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")

#---line plot
median_ysmooth_logcounts <- apply(logcounts_diff_mat,2,function(x){tapply(x,anno_row[names(x),"cluster"],mean)})
median_value_melt <- reshape2::melt(t(median_ysmooth_logcounts))
colnames(median_value_melt) <- c("pseudotime","cluster","logcounts")

library(ggplot2)
q <- ggplot(median_value_melt,aes(x=pseudotime,y=logcounts))+
  geom_smooth()+
 # scale_y_continuous(limits = c(0,8),breaks = seq(0,8,by=4))+
  facet_wrap(.~cluster,scales = "free_y",ncol=1)
q
ggsave(filename ="differential_all_genes_heatmap_cluster_line_for_TF_only.pdf" ,plot = q, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/figure/",
       width = 30, height =200, dpi = 600, limitsize = FALSE, units ="mm")
```


```{r}
logcounts_diff_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_all_diff_gene_frm.tsv",sep="\t",header=T)
logcounts_diff_frm$gene_name <- row.names(logcounts_diff_frm)

logcounts_diff_tf_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/ysmooth_logcounts_frm-V2.tsv",sep="\t",header=T)

logcounts_diff_gene_tf_overlap <- logcounts_diff_frm[row.names(logcounts_diff_tf_frm),]
logcounts_diff_gene_tf_overlap$tf_cluster <- logcounts_diff_tf_frm[,"cluster"]



logcounts_diff_list <- split(logcounts_diff_frm$gene_name,logcounts_diff_frm$cluster)

reactome_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/scATFR/inst/extdata/Reactome_pathway.rds")
mouse_reactome <- reactome_pathway$`Mus musculus`
mouse_reactome_num <- sapply(mouse_reactome,length)
mouse_reactome_filter <- mouse_reactome[between(mouse_reactome_num,10,500)]

score_cst_regulons_reactome <- calcuModuleSimilar(tfr_list = logcounts_diff_list, pathway_list = mouse_reactome_filter, adjust = "BH" , ncores=6, fdr = 0.05, min_overlap = 5)
saveRDS(score_cst_regulons_reactome,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/TF_logcounts_8_pattern_all_gene_pathway_enrichment.rds")
fwrite(score_cst_regulons_reactome,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/TF_logcounts_8_pattern_all_gene_pathway_enrichment.tsv",sep="\t")


```

#### 3.5 all three methods enrhchment comparison

```{r}
tf_exp_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/TF_logcounts_8_pattern_all_gene_pathway_enrichment.rds")
tf_regulon_pathway <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/All_8_patterns_reactome_enrichment_res.tsv",sep="\t",header=T,data.table = F)
bulk_tf_exp_pathway <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/EHT_bulk_seq_each_pattern_reactome_enrichment.tsv",sep="\t",header=T,data.table = F)


```

#### 3.6 selected pathway activity

```{r}
cts_tf_regulon_pathway <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/All_8_patterns_reactome_enrichment_res.rds")
head(sort(table(cts_tf_regulon_pathway$pathway),decreasing = T))
cluster_1_3_pathway <- cts_tf_regulon_pathway %>% filter(pattern %in% c("cluster1","cluster3"))
cell_cycle_pathway <- cluster_1_3_pathway %>% filter(pathway %in% c("Cell Cycle, Mitotic"))
splicing_pathway <- cluster_1_3_pathway %>% filter(pathway %in% c("Processing of Capped Intron-Containing Pre-mRNA"))
Autophagy_pathway <- cluster_1_3_pathway %>% filter(pathway %in% c("Autophagy"))

act_score_mat <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/smooth_value/ysmooth_active_score_tbl.rds")
act_score_diff <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,data.table=F)

cell_cycle_pathway_mat <- act_score_mat[intersect(act_score_diff$gene_name, splicing_pathway$regulon),]
splicing_pathway_mat <- act_score_mat[intersect(act_score_diff$gene_name, cell_cycle_pathway$regulon),]
Autophagy_pathway_mat <- act_score_mat[intersect(act_score_diff$gene_name, Autophagy_pathway$regulon),]

TFs_mat <- act_score_mat[c("Cebpa","Prdm16","Twist1"),]

plot_mat <- rbind(cell_cycle_pathway_mat,splicing_pathway_mat, Autophagy_pathway_mat, TFs_mat)
gene_name_rank <-c(row.names(cell_cycle_pathway_mat),row.names(splicing_pathway_mat),row.names(Autophagy_pathway_mat),c("Cebpa","Prdm16","Twist1"))
gene_name_rank_num <-c(nrow(cell_cycle_pathway_mat),nrow(splicing_pathway_mat),nrow(Autophagy_pathway_mat))

library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(9, "PRGn")
pal <- colorRampPalette(rev(cols))(100)
p <- pheatmap(mat = plot_mat,scale = "row",show_rownames = T,show_colnames = F,color = pal, silent = F, cluster_rows = F,cluster_cols = F,gaps_row = cumsum(gene_name_rank_num),labels_row = gene_name_rank, fontsize = 4)

ggsave(filename ="main_pathway_neg_gene_heatmap.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 100, height =100, dpi = 600, limitsize = FALSE, units ="mm")

TFs_mat <- act_score_mat[c("Cebpa","Prdm16","Twist1"),]
cols <- c("#1565c0","white","#c62828")
pal <- colorRampPalette(cols)(100)
p <- pheatmap(mat = TFs_mat,scale = "row",show_rownames = T,show_colnames = F,color = pal, silent = F, cluster_rows = F,cluster_cols = F)
ggsave(filename ="Cebpa_Prdm16_gene_act_heatmap.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 200, height =50, dpi = 600, limitsize = FALSE, units ="mm")



```





#### 4 KO data test

```{r}
Cebpa_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE42234.Cebpa.D21.table.tsv",sep="\t",header=T,data.table = F)
Cebpa_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE42234.Cebpa.table.tsv",sep="\t",header=T,data.table = F)

Cebpa_KO <- Cebpa_KO %>% tidyr::separate_rows(Gene.symbol,sep = '///')


cts_tf_regulon_df <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_diff_res.rds")
Cebpa_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Cebpa")

Cebpa_KO_logFC <- tapply(Cebpa_KO$logFC,Cebpa_KO$Gene.symbol,mean)
Cebpa_KO_logFC <- Cebpa_KO_logFC[-1]
library(fgsea)
enrich <- fgsea(Cebpa_regulon_df$targets, Cebpa_KO_logFC)
p <- plotEnrichment(Cebpa_regulon_df$targets[["Cebpa"]],Cebpa_KO_logFC)
p

ggsave(filename ="Cebpa_KO_diff.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/figure",
       width = 150, height =100, dpi = 600, limitsize = FALSE, units ="mm")

```

#### edger diff

```{r}
doedgeR <- function(x, group){
library(edgeR)
x = x
n = ncol(x)/2
group = group
y <- DGEList(counts = x, group = group)
#keep <- rowSums(cpm(y)>1) >= 1
#y <- y[keep, , keep.lib.sizes = FALSE]
y <- calcNormFactors(y)
design <- model.matrix(~group)
if(n == 1){
bcv <- 0.1
ltr <- exactTest(y,dispersion=bcv^2)
}else{
y <- estimateDisp(y,design)
fit <- glmFit(y,design)
ltr <- glmLRT(fit, coef=2)
}
FDR <- p.adjust(ltr$table$PValue, method="BH")
Direction <- decideTestsDGE(ltr,lfc=1)
res<-cbind(ltr$table,FDR,Direction)
res_order = res[order(res[,"FDR"]),]
return(res_order)
}

Prdm16_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE112860_VAVFL_RNAseq_Counts.txt.gz",sep="\t",header=T,data.table = F)
Prdm16_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE112860_47bpFL_RNAseq_Counts.txt.gz",sep="\t",header=T,data.table = F)
Prdm16_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE112860_VAVBM_RNAseq_Counts.txt/GSE112860_VAVBM_RNAseq_Counts.txt",sep="\t",header=T,data.table = F)
Prdm16_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE154011_Prdm16_normalized_gene_counts_matrix.txt.gz",sep="\t",header=T,data.table = F)
row.names(Prdm16_KO) <- Prdm16_KO$GeneID
gene_infor <- fread("D:/HSC_Autophagy/important_genes/orthlogo_gene/NCBI/gene_info/gene_info",sep="\t",header=T)
gene_infor <- gene_infor %>% filter(GeneID %in% Prdm16_KO$GeneID) %>% select(GeneID, Symbol)
Prdm16_KO <- inner_join(Prdm16_KO,gene_infor,by=c("GeneID"="GeneID"))

Prdm16_KO_mat <- apply(Prdm16_KO[,2:7],2,function(x){tapply(x,Prdm16_KO[,8],sum)})
Prdm16_KO_mat <- t(apply(Prdm16_KO_mat,1,function(x){(x-min(x))/(max(x)-min(x))}))
Prdm16_KO_logFC <- log2((apply(Prdm16_KO_mat[,4:6],1,mean))/(apply(Prdm16_KO_mat[,1:3],1,mean)))

Prdm16_KO_mat <- Prdm16_KO_mat[rowSums(Prdm16_KO_mat)>10,]
group <- colnames(Prdm16_KO_mat)
condition = factor(c(rep("WT",4),rep("KO",4)),levels=c("WT","KO"))

de_res = doedgeR(x = Prdm16_KO_mat, group = condition)
Prdm16_KO_logFC <- de_res[,"logFC"]
names(Prdm16_KO_logFC) <-row.names(de_res)

cts_tf_regulon_df <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_diff_res.rds")
Prdm16_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Prdm16")
Prdm16_regulon_list <- Prdm16_regulon_df$targets
Prdm16_regulon_list <- lapply(Prdm16_regulon_list, function(x){intersect(x,names(Prdm16_KO_logFC))})

library(fgsea)
enrich <- fgsea(Prdm16_regulon_list, Prdm16_KO_logFC)
p <- plotEnrichment(Prdm16_regulon_list$Prdm16,Prdm16_KO_logFC)
p
ggsave(filename ="Prdm16_KO_diff.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/figure",
       width = 150, height =100, dpi = 600, limitsize = FALSE, units ="mm")



```

#### 4.3 Twist1

```{r}
doedgeR <- function(x, group){
library(edgeR)
x = x
n = ncol(x)/2
group = group
y <- DGEList(counts = x, group = group)
#keep <- rowSums(cpm(y)>1) >= 1
#y <- y[keep, , keep.lib.sizes = FALSE]
y <- calcNormFactors(y)
design <- model.matrix(~group)
if(n == 1){
bcv <- 0.1
ltr <- exactTest(y,dispersion=bcv^2)
}else{
y <- estimateDisp(y,design)
fit <- glmFit(y,design)
ltr <- glmLRT(fit, coef=2)
}
FDR <- p.adjust(ltr$table$PValue, method="BH")
Direction <- decideTestsDGE(ltr,lfc=1)
res<-cbind(ltr$table,FDR,Direction)
res_order = res[order(res[,"FDR"]),]
return(res_order)
}
Twist1_WT <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE107814_RAW_Twist1/GSM2879636_WT.txt.gz",header=T,data.table = F)
Twist1_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/GSE107814_RAW_Twist1/GSM2879637_Twist1_KO.txt/GSM2879637_Twist1_KO.txt",header=T,data.table = F)
Twist1_KO_all <- inner_join(Twist1_WT,Twist1_KO,by=c('geneID'='geneID'))
gene_infor <- read.table("D:/HSC_Autophagy/ref_genome/gencode.vM25.gene.infor.rm.version.tsv",sep="\t",header=T,row.names=1)
Twist1_KO_all$gene_name <- gene_infor[Twist1_KO_all$geneID,]$gene_name
Twist1_KO_all$logFC <- log2(Twist1_KO_all$`Twist1 KO`/Twist1_KO_all$WT)
Twist1_KO_logFC <- tapply(Twist1_KO_all$logFC,Twist1_KO_all$gene_name,mean)
Twist1_KO_logFC <- scale(Twist1_KO_logFC)

cts_tf_regulon_df <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_diff_res.rds")
Twist1_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Twist1")
Twist1_regulon_list <- Twist1_regulon_df$targets[1]
#Prdm16_regulon_list <- lapply(Prdm16_regulon_list, function(x){intersect(x,names(Prdm16_KO_logFC))})
library(fgsea)
enrich <- fgsea(Twist1_regulon_list, Twist1_KO_logFC)
p <- plotEnrichment(Twist1_regulon_list$Twist1,Twist1_KO_logFC)
p

```

#### Tcf15

```{r}
Tcf15_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/Tcf15/GSE42539.D0.OE.table.tsv",sep="\t",header=T,data.table = F)
Tcf15_KO <- fread("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/Tcf15/GSE42539.D8.OE.table.tsv",sep="\t",header=T,data.table = F)
Tcf15_KO <- Tcf15_KO %>% tidyr::separate_rows(Gene.symbol,sep = '///')

cts_tf_regulon_df <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/diff_tfs/cts_tf_regulons_diff_res.rds")
Tcf15_regulon_df <- cts_tf_regulon_df %>% filter(regulon %in% "Tcf15")

Tcf15_KO_logFC <- tapply(Tcf15_KO$logFC,Tcf15_KO$Gene.symbol,mean)
Tcf15_KO_logFC <- Tcf15_KO_logFC[-1]
Tcf15_regulon_list <- lapply(Tcf15_regulon_df$targets,function(x){intersect(x,Tcf15_KO$Gene.symbol)})
library(fgsea)
enrich <- fgsea(Tcf15_regulon_list, Tcf15_KO_logFC)
p <- plotEnrichment(Tcf15_regulon_df$targets[["Tcf15"]],Tcf15_KO_logFC)
p

ggsave(filename ="Tcf15_OE_diff.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/test_KO_data/figure",
       width = 150, height =100, dpi = 600, limitsize = FALSE, units ="mm")
```



#### 5 regulon TFs DNA binding domain 

```{r}
activate_score_reactome_cluster1 <- score_cst_regulons_reactome %>% filter(pattern=="cluster1")
activate_score_reactome_others <- score_cst_regulons_reactome %>% filter(pattern!="cluster1")
activate_score_reactome_cluster1_only <- activate_score_reactome_cluster1 %>% filter(! pathway %in% activate_score_reactome_others$pathway) %>% filter(firstInGroup==1)
activate_score_reactome_cluster1_only






regulon_reactome_pathway <- regulon_reactome_pathway %>% filter(padj<0.05 & sapply(overlap_targets,length)>=5)
activate_score_reactome <- regulon_reactome_pathway %>% filter(regulon %in% row.names(activate_score_frm))
activate_score_reactome$cluster <- activate_score_frm[activate_score_reactome$regulon,"cluster"]
activate_score_reactome_cluster1 <- activate_score_reactome %>% filter(cluster=="cluster1")
activate_score_reactome_cluster1_mrege <- inner_join(activate_score_reactome_cluster1,activate_score_reactome_cluster1, by=c("regulon"="regulon"))
activate_score_reactome_cluster1_mrege <- activate_score_reactome_cluster1_mrege[,c("regulon","pathway.x","pathway.y")]
activate_score_reactome_cluster1_mrege <- activate_score_reactome_cluster1_mrege %>% filter(pathway.x != pathway.y)


activate_score_reactome_cluster1_pathway <- table(activate_score_reactome_cluster1$pathway)
activate_score_reactome_cluster1_pathway_2 <- activate_score_reactome_cluster1_pathway[activate_score_reactome_cluster1_pathway>1]

TF_infors <- fread("D:/HSC_Autophagy/important_genes/Human_TFs/TF_infors_v_1.01.txt",sep="\t",header=T)
TF_infors <- TF_infors[,3:4]
TF_infors_mmu <- getHomologGene(input_gene = TF_infors$`HGNC symbol`,output_type = "gene_symbol")
TF_infors_mmu <- inner_join(TF_infors_mmu,TF_infors,by=c('Gene_Symbol_human'='HGNC symbol'))
fwrite(TF_infors_mmu,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/TFs/hsa_mmu_TF_domain_infor.tsv",sep="\t")
TF_infors_mmu_filter <- TF_infors_mmu[,-1] %>% unique() %>% as.data.frame()
row.names(TF_infors_mmu_filter) <- TF_infors_mmu_filter$Gene_Symbol_mouse
activate_score_frm$DBD <- TF_infors_mmu_filter[row.names(activate_score_frm),"DBD"]
activate_score_frm$gene_name <- row.names(activate_score_frm)
activate_score_frm_new <- activate_score_frm %>% tidyr::separate_rows(DBD,sep="; ")


stat <- table(activate_score_frm_new$DBD,activate_score_frm_new$cluster)
stat_filter <- stat[rowSums(stat)>=1,]
stat_ratio <- apply(stat_filter,2,function(x){x/sum(x)})

library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "YlOrRd")
cols <- c("#aa00ff","black","#ffff00")
pal <- colorRampPalette(cols)(100)
anno_row <- as.data.frame(log(rowSums(stat)+1))
colnames(anno_row) <- "num"
p <- pheatmap(t(stat_ratio), scale="column", cluster_rows = F, clustering_method = "ward.D2", angle_col=315, color = pal, cutree_cols = 8, fontsize = 8,annotation_col = anno_row)
ggsave(filename ="pattern_domain_heatmap.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/figure/",
       width = 250, height =70, dpi = 600, limitsize = FALSE, units ="mm")


fwrite(activate_score_reactome_cluster1_mrege,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/cluster1.tsv",sep="\t")


activate_score_reactome_cluster2 <- activate_score_reactome %>% filter(cluster=="cluster2")
activate_score_reactome_cluster3 <- activate_score_reactome %>% filter(cluster=="cluster3")
activate_score_reactome_cluster4 <- activate_score_reactome %>% filter(cluster=="cluster4")
activate_score_reactome_cluster5 <- activate_score_reactome %>% filter(cluster=="cluster5")
activate_score_reactome_cluster6 <- activate_score_reactome %>% filter(cluster=="cluster6")
activate_score_reactome_cluster7 <- activate_score_reactome %>% filter(cluster=="cluster7")
activate_score_reactome_cluster8 <- activate_score_reactome_ %>% filter(cluster=="cluster8")

#--autophagy

autophagy <- regulon_reactome_pathway %>% filter(pathway %in% "Autophagy")
autophagy$cluster <- activate_score_frm[autophagy$regulon,"cluster"]
autophagy_regulons <- na.omit(autophagy)
```


#### 2.4 splicing pathway analysis
```{r}
pattern_reactome <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/All_8_patterns_reactome_enrichment_res.rds")

splicing_pathway <- pattern_reactome %>% filter(pathway %in% 'mRNA Splicing')
splicing_pathway_genes <- splicing_pathway %>% select(regulon,overlap_targets,pattern)
splicing_pathway_TF_Target <- purrr::map2(splicing_pathway_genes$overlap_targets, splicing_pathway_genes$regulon,
                                          .f=function(x,y){
                                            data.frame(tf=rep(y,length(x)),target=x)
                                          })
splicing_pathway_TF_Target <- do.call(rbind,splicing_pathway_TF_Target) %>% filter(tf != target) %>% unique()
splicing_pathway_TF_Target$pattern <- activate_score_frm[splicing_pathway_TF_Target$tf,"cluster"]
fwrite(splicing_pathway_TF_Target,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/mRNA_splicing.tsv",sep="\t")
splicing_pathway_TFs <- unique(splicing_pathway_TF_Target$tf)
splicing_pathway_targets <- setdiff(unique(splicing_pathway_TF_Target$target),splicing_pathway_TFs)
tf_clusters <- unique(splicing_pathway_TF_Target[,c(1,3)])
colnames(tf_clusters) <- c("gene","cluster")
target_clusters <- unique(splicing_pathway_TF_Target[,-1])
colnames(target_clusters) <- c("gene","cluster")
tf_target_clusters <- rbind(tf_clusters, target_clusters) %>% unique()
tf_target_clusters_attr <- tapply(tf_target_clusters$cluster,tf_target_clusters$gene,paste0,collapse=";")

splicing_pathway_attritube <- data.frame(gene =c(splicing_pathway_TFs,splicing_pathway_targets),
                                         attr = c(rep("tf",length(splicing_pathway_TFs)), 
                                                  rep("target",length(splicing_pathway_targets))))
splicing_pathway_attritube$attr2 <- tf_target_clusters_attr[splicing_pathway_attritube$gene]

fwrite(splicing_pathway_attritube,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/pathway/mRNA_splicing_node_attritube.tsv",sep="\t")

activate_score_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/EHT_analysis/pseudotime/diff_res/activate_score_frm.tsv",sep="\t",header=T,row.names=1)

splicing_DBD <- activate_score_frm_new %>% filter(gene_name %in% splicing_pathway_TFs)

```
 
#### 2.4 autophagy pathway analysis

```{r}

```


#### 2.2 divergent TF and TF regulons according pseudotime

```{r}


```

