---
title: "R Notebook"
output: html_notebook
---

#### 1. regulon performance test using different regulon sources

#### 1.1 load_data

```{r}
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/objects/atfr.rds")
col_data <- colData(atfr)
TFs <- setdiff(unique(col_data$TF),"Scramble")

```

###$ 1.2 performance test

```{r}
get_pref <- function(mat, col_data, grn_method="pidc", act_method="viper", filter=TRUE){
  cells <- colnames(mat)
  row.names(col_data) <- col_data$cell
  labels <- col_data[cells,"TF"]
  TF_gene <- setdiff(labels,"Scramble")
  names(labels) <- cells
  label <- ifelse(labels=="Scramble",1,-1)
  prediction <- mat[1,names(labels)]
  pred <- ROCR::prediction(predictions = prediction, labels = label)
  perf <- ROCR::performance(pred,"auc")
  auc_value <- perf@y.values[[1]]
  perf <- ROCR::performance(pred,"tpr", "fpr")
  out_res <- data.frame(TF=rep(TF_gene,length(perf@x.values[[1]])),
                        FPR = perf@x.values[[1]],
                        TPR = perf@y.values[[1]],
                        AUC = rep(auc_value,length(perf@x.values[[1]])),
                        GRNs = rep(grn_method,length(perf@x.values[[1]])),
                        Activity = rep(act_method, length(perf@x.values[[1]])),
                        withFilter = rep(filter,length(perf@x.values[[1]])))
  out_res
}

get_pref_mat <- function(x, i="viper",...){
  act_mat <- assay(altExp(atfr,i))
  TFs <- setdiff(row.names(act_mat),'Scramble')
  col_data <- colData(x)
  row.names(col_data) <- col_data$cells
  .get_mat <- function(tf,mat,col_data){
    case_cells <- col_data[col_data$TF %in% tf,"cell"]
    control_cells <- col_data[col_data$TF %in% 'Scramble',"cell"]
    mat[tf,c(control_cells,case_cells),drop=FALSE]
  }
  act_mat_list <- lapply(TFs,.get_mat,mat=act_mat,col_data=col_data)
  names(act_mat_list) <- TFs
  act_pref_list <- lapply(act_mat_list,get_pref,act_method=i,col_data=col_data,...)
  act_pref_list
}

#---pidc
viper_pref <- get_pref_mat(x = atfr,grn_method="pidc",i = "viper")
viper_pref <- do.call(rbind,viper_pref)
viper_TF_auc <- unique(viper_pref[,c('TF','AUC',"Activity")])
row.names(viper_TF_auc) <- viper_TF_auc$TF
```



#### 1.3 integration vs no integration

#### 1.3.1 merged pantissue network
```{r}
pantissue_network <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/final_network/human/pantissue_merged_filter_strict_network.txt.gz",sep="\t",header=T)
pantissue_network_filter <- pantissue_network %>% filter(tf %in% TFs)
PIDC_GRN_aracne <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/GRNs/PIDC_GRN_regulon.rds")
PIDC_aracne_regulon <- PIDC_GRN_aracne %>% filter(regulator %in% TFs)
colnames(PIDC_aracne_regulon) <- c("tf","target","weight")

merge_PIDC_nets <- integraNets(grnNets = list(PIDC_aracne_regulon, pantissue_network_filter), join_type = "full", norm_weight =FALSE, return_sig = T, min_target_num = 10, max_target_num = 2000, ncores = 6, maxInter = 5)
merge_PIDC_regulons <- split(merge_PIDC_nets$target,merge_PIDC_nets$tf)

atfr <- regulonActivity(atfr, method="viper", gene_list=merge_PIDC_regulons, use_assay="logcounts", ncores=1,max_target_num=1000L)
merged_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
merged_pref <- do.call(rbind, merged_pref)
merged_pref <- unique(merged_pref[,c(1,4)])
colnames(merged_pref)[2] <- "merged"
#atfr <- regulonActivity(atfr, method="ssgsea", gene_list=merge_PIDC_regulons, use_assay="logcounts", ncores=1, max_target_num=1000L)
#ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc",i = "ssgsea")
#ssgsea_pref <- do.call(rbind,ssgsea_pref)
#ssgsea_pref <- unique(ssgsea_pref[,c(1,4,6)])

#merged_regulon_AUC <- list(viper=viper_pref,ssgsea=ssgsea_pref)

```

#### 1.3.2 expression network only
```{r}
PIDC_GRN_aracne <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/GRNs/PIDC_GRN_regulon.rds")
PIDC_GRN_aracne <- PIDC_GRN_aracne %>% filter(regulator %in% TFs)
PIDC_GRN_aracne_regulon <- split(as.character(PIDC_GRN_aracne$target), as.character(PIDC_GRN_aracne$regulator))

atfr <- regulonActivity(atfr, method="viper", gene_list=PIDC_GRN_aracne_regulon, use_assay="logcounts", ncores=1, max_target_num=2000L)
aracne_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
aracne_pref <- do.call(rbind, aracne_pref)
aracne_pref <- unique(aracne_pref[,c(1,4)])
colnames(aracne_pref)[2] <- "aracne"

#atfr <- regulonActivity(atfr, method="ssgsea", gene_list=PIDC_GRN_aracne_regulon, use_assay="logcounts", ncores=1, max_target_num=1000L)
#ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc",i = "ssgsea")
#ssgsea_pref <- do.call(rbind,ssgsea_pref)
#ssgsea_pref <- unique(ssgsea_pref[,c(1,4,6)])
#aracne_regulon_AUC <- list(viper=viper_pref,ssgsea=ssgsea_pref)

```

#### 1.3.3 integered prior network only

```{r}
pantissue_network <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/final_network/human/pantissue_merged_filter_strict_network.txt.gz",sep="\t",header=T)
pantissue_network_filter <- pantissue_network %>% filter(tf %in% TFs)
pantissue_regulon <- split(as.character(pantissue_network_filter$target),as.character(pantissue_network_filter$tf))

atfr <- regulonActivity(atfr, method="viper", gene_list=pantissue_regulon, use_assay="logcounts", ncores=1, max_target_num=2000L)
pantissue_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
pantissue_pref <- do.call(rbind, pantissue_pref)
pantissue_pref <- unique(pantissue_pref[,c(1,4)])
colnames(pantissue_pref)[2] <- "pantissue"

#atfr <- regulonActivity(atfr, method="ssgsea", gene_list=pantissue_regulon, use_assay="logcounts", ncores=1, max_target_num=2000L)
#ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc",i = "ssgsea")
#ssgsea_pref <- do.call(rbind,ssgsea_pref)
#ssgsea_pref <- unique(ssgsea_pref[,c(1,4,6)])
#pantissue_regulon_AUC <- list(viper=viper_pref, ssgsea=ssgsea_pref)

```

#### 1.3.4 prior network from dorothea

```{r}
dorothea_regulons <- dorothea::dorothea_hs %>% filter(tf %in% TFs)
dorothea_regulon_list <- split(dorothea_regulons$target,dorothea_regulons$tf)

atfr <- regulonActivity(atfr, method="viper", gene_list=dorothea_regulon_list, use_assay="logcounts", ncores=1, max_target_num=5000L, min_target_num=5L)
dorothea_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
dorothea_pref <- do.call(rbind, dorothea_pref)
dorothea_pref <- unique(dorothea_pref[,c(1,4)])
colnames(dorothea_pref)[2] <- "dorothea"
```


#### 1.3.5 prior network from SCENIC

```{r}
scenic_regulons <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/SCENIC_workflow/grn_regulon_list.rds")

atfr <- regulonActivity(atfr, method="viper", gene_list=scenic_regulons, use_assay="logcounts", ncores=1, max_target_num=1000L, min_target_num=5L)
scenic_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
scenic_pref <- do.call(rbind, scenic_pref)
scenic_pref <- unique(scenic_pref[,c(1,4)])
colnames(scenic_pref)[2] <- "SCENIC"


```


#### 1.3.6 all regulons performance plot

```{r}
all_pref_list <- list(merged_pref,aracne_pref,pantissue_pref,dorothea_pref,scenic_pref)
all_pref_frm <- purrr::reduce(.x = all_pref_list, .f = left_join,by=c("TF"="TF"))
fwrite(all_pref_frm,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/5_regulon_source_AUC_value.tsv",sep="\t")
all_pref_frm_melt <- reshape2::melt(all_pref_frm)
colnames(all_pref_frm_melt)[2:3] <- c("method","AUROC")
plot_frm <- all_pref_frm_melt
library(ggplot2)
text_size = 8
line_size = 0.5
mytheme <- theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA),
panel.grid.major = element_line(size=0.25*line_size, linetype = 8, color = "lightgray"),
axis.ticks = element_line(size=line_size),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45,hjust=1,vjust=1,color="black"), 
axis.text.y= element_text(size=text_size, color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

plot_frm$method <- factor(plot_frm$method,levels=c("dorothea","SCENIC","pantissue","aracne","merged"))

p <- ggplot(data = plot_frm, aes(x = method, y=AUROC, fill = method))+
  geom_violin(color="black")+
#  geom_jitter(width=0.1)+
  geom_boxplot(width=0.15,fill="white",size=0.5,outlier.color = NA)+
  mytheme+
  scale_y_continuous(limits = c(0.35,1))
p

ggsave(filename = "merged_regulons_AUROC_comparison.pdf", plot = p, device = "pdf", path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/figure/", width = 100, height = 70, units = "mm")


plot_density <- plot_frm
plot_density$method <- factor(plot_density$method, levels = rev(levels(plot_density$method)))
library(ggridges)
q <- ggplot(data = plot_density, aes(x = AUROC, y = method, fill = method, height=..density..))+
  geom_density_ridges_gradient(size = line_size, scale = 1.4, panel_scaling = F, alpha=.8,na.rm=T)+
  mytheme+
#  coord_flip()+
  scale_x_continuous(limits = c(0.4,1))
q

#----significant test
row.names(all_pref_frm) <- all_pref_frm$TF
sources <- c("dorothea","SCENIC","pantissue","aracne","merged")
test_mat <- matrix(1,nrow = 5,ncol=5)
dimnames(test_mat) <- list(sources,sources)
for(i in sources){
  for(j in setdiff(sources,i)){
    test_val <- all_pref_frm[,c(i,j)]
    test_val <- na.omit(test_val)
    test_mat[i,j] <- t.test(test_val[,i],test_val[,j],paired = T)$p.value
  }
}

cormat <- -log10(test_mat)
library(pheatmap)
library(RColorBrewer)
cols <- brewer.pal(3, "PRGn")
pal <- colorRampPalette(rev(cols))(100)
p <- pheatmap(cormat,cluster_rows = T,cluster_cols = T,scale="none",display_numbers = T,color = pal, breaks = seq(0,4,length.out=100), number_format = "%.1f", border_color = NA, legend_breaks = c(0,1.30103,2,4))
ggsave(filename ="5_regulon_source_paired_t_test_pval.pdf" ,plot = p, device = "pdf", 
       path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/figure/",
       width = 100, height =85, dpi = 600, limitsize = FALSE, units ="mm")


```


```{r}
geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                        position = "dodge", trim = TRUE, scale = "area",
                        show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)

            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(ymin = min(y),
                     ymax = max(y),
                     xmin = x - width / 2,
                     xmax = x)
          },

          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data, 
                              xmaxv = x,
                              xminv = x + violinwidth * (xmin - x))

            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(plyr::arrange(transform(data, x = xminv), y),
                             plyr::arrange(transform(data, x = xmaxv), -y))

            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1,])

            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },

          draw_key = draw_key_polygon,

          default_aes = aes(weight = 1, colour = "grey20", fill = "white", size = 0.5,
                            alpha = NA, linetype = "solid"),

          required_aes = c("x", "y")
)
```

#### 2 AUcell ssgsea viper performance


#### 2.1 regulons from integrated source 

```{r}
all_pref_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/5_regulon_source_AUC_value.tsv",sep="\t",header=T)
row.names(all_pref_frm) <- all_pref_frm$TF
viper_pref <- all_pref_frm[,1:2]
colnames(viper_pref)[2] <- "viper"

#---
atfr <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/objects/atfr.rds")
col_data <- colData(atfr)
TFs <- setdiff(unique(col_data$TF),"Scramble")

pantissue_network <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/TF_Target_data/final_network/human/pantissue_merged_filter_strict_network.txt.gz",sep="\t",header=T)
pantissue_network_filter <- pantissue_network %>% filter(tf %in% TFs)
PIDC_GRN_aracne <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/GRNs/PIDC_GRN_regulon.rds")
PIDC_aracne_regulon <- PIDC_GRN_aracne %>% filter(regulator %in% TFs)
colnames(PIDC_aracne_regulon) <- c("tf","target","weight")

merge_PIDC_nets <- integraNets(grnNets = list(PIDC_aracne_regulon, pantissue_network_filter), join_type = "full", norm_weight =FALSE, return_sig = T, min_target_num = 10, max_target_num = 2000, ncores = 2, maxInter = 5)
merge_PIDC_regulons <- split(merge_PIDC_nets$target,merge_PIDC_nets$tf)

atfr <- regulonActivity(atfr, method="ssgsea", gene_list=merge_PIDC_regulons, use_assay="logcounts", ncores=1,max_target_num=1000L)
ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "ssgsea")
ssgsea_pref <- do.call(rbind,ssgsea_pref)
ssgsea_pref_auc <- unique(ssgsea_pref[,c("TF","AUC")])
row.names(ssgsea_pref_auc) <- ssgsea_pref_auc$TF
colnames(ssgsea_pref_auc)[2] <- "ssgsea"

atfr <- regulonActivity(atfr, method="aucell", gene_list=merge_PIDC_regulons, use_assay="logcounts", ncores=1, max_target_num=1000L)
aucell_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "aucell")
aucell_pref <- do.call(rbind,aucell_pref)
aucell_pref_auc <- unique(aucell_pref[,c("TF","AUC")])
row.names(aucell_pref_auc) <- aucell_pref_auc$TF
colnames(aucell_pref_auc)[2] <- "aucell"

scATFR_pref <- purrr::reduce(.x = list(viper_pref, ssgsea_pref_auc, aucell_pref_auc),.f=left_join,by=c("TF"="TF"))
saveRDS(scATFR_pref,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/activate_test/scATFR_act_pref_auc.rds")

scATFR_pref_melt <- reshape2::melt(scATFR_pref)
colnames(scATFR_pref_melt)[2:3] <- c("method","AUC")
scATFR_pref_melt$software <- rep("scATFR",nrow(scATFR_pref_melt))

t.test(scATFR_pref$viper,scATFR_pref$aucell,paired = T)
t.test(scATFR_pref$ssgsea,scATFR_pref$aucell,paired = T)
t.test(scATFR_pref$ssgsea,scATFR_pref$viper,paired = T)

```

#### 2.2 regulons from SCENIC


```{r}
rcistarget_regulon <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/SCENIC_workflow/grn_regulon_list.rds")
atfr <- regulonActivity(atfr, method="viper", gene_list=rcistarget_regulon, use_assay="logcounts", ncores=1,max_target_num=1000L)
scenic_viper_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
scenic_viper_pref <- do.call(rbind, scenic_viper_pref)
scenic_viper_pref_auc <- unique(scenic_viper_pref[,c("TF","AUC")])
colnames(scenic_viper_pref_auc)[2] <- "viper"

atfr <- regulonActivity(atfr, method="ssgsea", gene_list=rcistarget_regulon, use_assay="logcounts", ncores=1,max_target_num=1000L)
scenic_ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "ssgsea")
scenic_ssgsea_pref <- do.call(rbind, scenic_ssgsea_pref)
scenic_ssgsea_pref_auc <- unique(scenic_ssgsea_pref[,c("TF","AUC")])
colnames(scenic_ssgsea_pref_auc)[2] <- "ssgsea"

atfr <- regulonActivity(atfr, method="aucell", gene_list=rcistarget_regulon, use_assay="logcounts", ncores=1,max_target_num=1000L)
scenic_aucell_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "aucell")
scenic_aucell_pref <- do.call(rbind, scenic_aucell_pref)
scenic_aucell_pref_auc <- unique(scenic_aucell_pref[,c("TF","AUC")])
colnames(scenic_aucell_pref_auc)[2] <- "aucell"

scenic_pref <- purrr::reduce(.x = list(scenic_viper_pref_auc, scenic_ssgsea_pref_auc, scenic_aucell_pref_auc), .f=left_join, by=c("TF"="TF"))
saveRDS(scenic_pref,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/activate_test/SCENIC_act_pref_auc.rds")
scenic_pref_melt <- reshape2::melt(scenic_pref)
colnames(scenic_pref_melt)[2:3] <- c("method","AUC")
scenic_pref_melt$software <- rep("rcistarget",nrow(scenic_pref_melt))

```


### 2.3 dorothea regulons

```{r}
dorothea_regulons <- dorothea::dorothea_hs %>% filter(tf %in% TFs)
dorothea_regulon <- split(dorothea_regulons$target,dorothea_regulons$tf)
#saveRDS(dorothea_regulons,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/dorothea_workflow/dorothea_regulons.rds")

atfr <- regulonActivity(atfr, method="viper", gene_list=dorothea_regulon, use_assay="logcounts", ncores=1,max_target_num=2000L)
dorothea_viper_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
dorothea_viper_pref <- do.call(rbind, dorothea_viper_pref)
dorothea_viper_pref_auc <- unique(dorothea_viper_pref[,c("TF","AUC")])
colnames(dorothea_viper_pref_auc)[2] <- "viper"

atfr <- regulonActivity(atfr, method="ssgsea", gene_list=dorothea_regulon, use_assay="logcounts", ncores=1,max_target_num=2000L)
dorothea_ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "ssgsea")
dorothea_ssgsea_pref <- do.call(rbind, dorothea_ssgsea_pref)
dorothea_ssgsea_pref_auc <- unique(dorothea_ssgsea_pref[,c("TF","AUC")])
colnames(dorothea_ssgsea_pref_auc)[2] <- "ssgsea"

atfr <- regulonActivity(atfr, method="aucell", gene_list=dorothea_regulon, use_assay="logcounts", ncores=1,max_target_num=2000L)
dorothea_aucell_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "aucell")
dorothea_aucell_pref <- do.call(rbind, dorothea_aucell_pref)
dorothea_aucell_pref_auc <- unique(dorothea_aucell_pref[,c("TF","AUC")])
colnames(dorothea_aucell_pref_auc)[2] <- "aucell"

dorothea_pref <- purrr::reduce(.x = list(dorothea_viper_pref_auc, dorothea_ssgsea_pref_auc, dorothea_aucell_pref_auc), .f=left_join, by=c("TF"="TF"))
saveRDS(dorothea_pref,"D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/activate_test/dorothea_act_pref_auc.rds")
dorothea_pref_melt <- reshape2::melt(dorothea_pref)
colnames(dorothea_pref_melt)[2:3] <- c("method","AUC")
dorothea_pref_melt$software <- rep("dorothea",nrow(dorothea_pref_melt))

```

#### 2.4 merged table AUC

```{r}
all_pref_auc <- rbind(scATFR_pref_melt, scenic_pref_melt, dorothea_pref_melt, make.row.names =F)

plot_frm <- all_pref_auc
library(ggplot2)
text_size = 8
line_size = 0.5
mytheme <- theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA),
panel.grid.major = element_line(size=0.25*line_size, linetype = 8, color = "lightgray"),
axis.ticks = element_line(size=line_size),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45,hjust=1,vjust=1,color="black"), 
axis.text.y= element_text(size=text_size, color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

plot_frm$method <- factor(plot_frm$method,levels=c("dorothea","SCENIC","pantissue","aracne","merged"))

p <- ggplot(data = plot_frm, aes(x = method, y=AUC, fill = method))+
#  geom_violin(color="black")+
  geom_boxplot(width=0.4,size=0.5,outlier.color = NA)+
  geom_jitter(width=0.1,size=0.3)+
  mytheme+
  scale_y_continuous(limits = c(0.35,1))+
  facet_wrap(.~software,nrow=1)
p


plot_density <- plot_frm
plot_density$method <- factor(plot_density$method, levels = rev(levels(plot_density$method)))
library(ggridges)
q <- ggplot(data = plot_density, aes(x = AUC, y = method, fill = method, height=..density..))+
  geom_density_ridges_gradient(size = line_size, scale = 1.4, panel_scaling = F, alpha=.8,na.rm=T)+
  mytheme+
#  coord_flip()+
  scale_x_continuous(limits = c(0.4,1))+
  facet_wrap(.~software,nrow=1)
q
ggsave(filename = "viper_ssgsea_aucell_SCENIC_dorothea_scatfr_comparison_AUROC.pdf", plot = q, device = "pdf", path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/figure/", width = 150, height = 50, units = "mm")


```






#### 2.4 dorothea + gsea

```{r}

atfr <- filterRegulons(x = atfr,gene_list=dorothea_regulon, use_grn ="pidc", score_type="pos", maxSize = 1000L)
dorothea_regulon_filter <- as.list(Regulon(atfr))
atfr <- regulonActivity(atfr, method="viper", gene_list=dorothea_regulon_filter, use_assay="logcounts", ncores=1,max_target_num=1000L)
dorothea_viper_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "viper")
dorothea_viper_pref <- do.call(rbind, dorothea_viper_pref)
dorothea_viper_pref_auc <- unique(dorothea_viper_pref[,c("TF","AUC")])
colnames(dorothea_viper_pref_auc)[2] <- "viper"

atfr <- regulonActivity(atfr, method="ssgsea", gene_list=dorothea_regulon_filter, use_assay="logcounts", ncores=1,max_target_num=2000L)
dorothea_ssgsea_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "ssgsea")
dorothea_ssgsea_pref <- do.call(rbind, dorothea_ssgsea_pref)
dorothea_ssgsea_pref_auc <- unique(dorothea_ssgsea_pref[,c("TF","AUC")])
colnames(dorothea_ssgsea_pref_auc)[2] <- "ssgsea"

atfr <- regulonActivity(atfr, method="aucell", gene_list=dorothea_regulon_filter, use_assay="logcounts", ncores=1,max_target_num=2000L)
dorothea_aucell_pref <- get_pref_mat(x = atfr, grn_method="pidc", i = "aucell")
dorothea_aucell_pref <- do.call(rbind, dorothea_aucell_pref)
dorothea_aucell_pref_auc <- unique(dorothea_aucell_pref[,c("TF","AUC")])
colnames(dorothea_aucell_pref_auc)[2] <- "aucell"
dorothea_filter_pref <- purrr::reduce(.x = list(dorothea_viper_pref_auc, dorothea_ssgsea_pref_auc, dorothea_aucell_pref_auc), .f=left_join, by=c("TF"="TF"))


```



#### 3. software performance test using three different softwares.

#### 3.1 scATFR vs SCENIC vs dorothea

```{r}

all_pref_frm <- read.table("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/5_regulon_source_AUC_value.tsv",sep="\t",header=T)
SCENIC_pref <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/SCENIC_workflow/each_TF_AUC_value.rds")
dorothea_pref <- readRDS("D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/dorothea_workflow/ROC_AUC_perform.rds")

all_pref_frm_new <- all_pref_frm[,1:2]
row.names(all_pref_frm_new) <- all_pref_frm_new$TF
all_pref_frm_new$SCENIC <- SCENIC_pref[all_pref_frm_new$TF]
all_pref_frm_new$dorothea <- dorothea_pref[all_pref_frm_new$TF]

TF_AUC_melt <- reshape2::melt(all_pref_frm_new)
colnames(TF_AUC_melt)[2:3] <- c("method","AUROC")
ALL_TF_order <- all_pref_frm_new
ALL_TF_order$diff1 <- ALL_TF_order$integrated-ALL_TF_order$SCENIC
ALL_TF_order$diff2 <- ALL_TF_order$integrated-ALL_TF_order$dorothea
ALL_TF_order$diff <- ALL_TF_order$diff1+ALL_TF_order$diff2
ALL_TF_order <- ALL_TF_order[order(ALL_TF_order$diff,ALL_TF_order$diff1,ALL_TF_order$diff2),]
TF_AUC_melt$TF <- factor(TF_AUC_melt$TF,levels=ALL_TF_order$TF)

plot_frm <- TF_AUC_melt
library(ggplot2)
text_size = 8
line_size = 0.5
mytheme <- theme(line = element_line(size=line_size), text = element_text(size=text_size, color="black"), 
panel.border = element_rect(size = line_size, fill=NA),
panel.background = element_rect(fill=NA),
panel.grid.major = element_line(size=0.25*line_size, linetype = 8, color = "lightgray"),
axis.ticks = element_line(size=line_size),
axis.title.x = element_text(size=text_size),
axis.text.x= element_text(size=text_size, angle=45,hjust=1,vjust=1,color="black"), 
axis.text.y= element_text(size=text_size, color="black"),
legend.text = element_text(size = text_size), 
legend.title = element_text(size=text_size*1.2),
plot.title=element_text(size=text_size, hjust=0.5,vjust=0.5),legend.position='right')

p <- ggplot(plot_frm,aes(x=TF,y=AUROC,group=method,color=method))+
  geom_line()+
  geom_point(shape=16)+
  mytheme+
  scale_y_continuous(limits = c(0.4,1))
p

ggsave(filename = "PIDC_SCENIC_dorothea_main_comparison_AUROC.pdf", plot = p, device = "pdf", path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/figure/", width = 180, height = 50, units = "mm")



plot_density <- plot_frm
plot_density$method <- factor(plot_density$method, levels = rev(levels(plot_density$method)))
library(ggridges)
q <- ggplot(data = plot_density, aes(x = AUROC, y = method, fill = method, height=..density..))+
  geom_density_ridges_gradient(size = line_size, scale = 1.4, panel_scaling = F, alpha=.8,na.rm=T)+
  mytheme+
  coord_flip()+
  scale_x_continuous(limits = c(0.4,1))
q
ggsave(filename = "PIDC_SCENIC_dorothea_main_comparison_AUROC_density.pdf", plot = q, device = "pdf", path = "D:/HSC_Autophagy/Markdown/3.scATFR/ext_data/CRISPRi/performance/figure/", width = 70, height = 50, units = "mm")


```








